\solution{\ref{problemConvergencesum_2^infty1/(xlnx)dx}.
The function $\frac{1}{x\ln x}$ is decreasing, as for $x>2$, it is the quotient of $1$ by increasing positive functions. $\frac{ 1}{ x\ln x}$ tends to $0$ as $x\to \infty$, and therefore the integral criterion implies that $\sum\limits_{n=2}^{ \infty} \frac{1 }{n \ln n}$ is convergent/divergent if and only if $\int_{2}^\infty \frac{1}{x\ln x}\diff x$ is.

\[
\begin{array}{rcl}
\displaystyle \int_{2}^{\infty}\frac{1}{x\ln x}\diff x&=&\displaystyle \lim_{t\to\infty} \int_{2}^{t}\frac{1}{x\ln x}\diff x\\
&=&\displaystyle \lim_{t\to\infty} \int_{2}^{t}\frac{1}{\ln x}\diff(\ln x)\\
&=&\displaystyle \lim_{t\to\infty} \int_{2}^{t}\diff(\ln(\ln x))\\
&=&\displaystyle \lim_{t\to \infty}\left[ \ln(\ln x)\right]_{x=2}^{x=t}\\
&=&\displaystyle \lim_{t\to \infty}\left(\ln(\ln t)-\ln (\ln 2)\right)\\
&=&\displaystyle \infty.
\end{array}
\]
The integral is divergent (and diverges to $+\infty$) and therefore, by the integral criterion, so is the sum.
}

\solution{\ref{problemConvergencesum_2^infty1/((2n+1)ln(n)}
The integral criterion appears to be of little help: the improper integral $\int \frac{1 }{(2x+1)\ln x}\diff x $ cannot be integrated algebraically with any of the techniques we have studied so far. Therefore it makes sense to try to solve this problem using a comparison test. 

We present two solution variants. In Variant I we use the limit-comparison test. This is an easier (but slightly longer) solution. In Variant II we use the comparison test - this solution is harder as it requires algebraic intuition to select a series to compare to.


\textbf{Variant I.} This variant uses the limit comparison test.  

The ``dominant term''\footnote{since we do not speak of rational functions, here the expression ``dominant term'' is used informally} of the denominator of $\frac{1}{(2n+1)\ln n}=\frac{1}{2n\ln n +\ln n}$ is $ 2n\ln n$. Therefore it makes sense to compare - or limit-compare - with $\frac{1}{n\ln n } $. 

We will use the Limit Comparison Test for the series $\sum\limits_{n=2}^\infty a_n = \sum\limits_{n=2}^\infty \frac{1}{(2n+1)\ln n}$ and $\sum\limits_{n=2}^{\infty} b_n= \sum\limits_{n=2}^\infty \frac{1}{n\ln n} $. Both $a_n$ and $b_n$ are positive (for $n>2$) and therefore the Limit Comparison Test applies.

\[
\lim\limits_{n\to \infty}\frac{a_n}{b_n} = \lim\limits_{n\to\infty} \frac{ \frac{1}{(2n+1)\ln n}}{\frac{1}{n\ln n}} = \lim\limits_{n\to\infty}\frac{ n}{2n+1}=\lim\limits_{n\to\infty}\frac{ 1}{2+\frac{1}{n}} = \frac{1}{2}.
\]
Since $\lim\limits_{n\to \infty}\frac{a_n}{b_n}=\frac{1}{2}\neq 0$, the Limit Comparison Test implies that the series $\sum\limits_{n=2}^{\infty} a_n$ has same convergence/divergence properties as the series $\sum\limits_{n=2}^{\infty} b_n$. In Problem \ref{problemConvergencesum_2^infty1/(xlnx)dx} we demonstrated that the series $\sum\limits_{n=2}^{\infty} b_n$ is divergent; therefore the series $\sum\limits_{n=2}^\infty a_n=  \sum \limits_{n=2 }^\infty \frac{1}{(2n+1)\ln n}$ is divergent as well.

\textbf{Variant II. } This variant uses directly the comparison test. It is slightly shorter than the preceding variant but requires more intuition.

Let $a_n= \frac{1}{(2n+1)\ln n}$. Consider the series $\sum\limits_{n=2}^{\infty}b_n$ for  $b_n=\frac{1}{3n \ln n}$. We have that 
\[
\begin{array}{rcll|l}
3n &\geq & 2n+1 &&\text{for } n\geq 1\\
\displaystyle \frac{1}{3n} &\leq&\displaystyle  \frac{1}{2n+1} \quad .&& \begin{array}{@{}l}\text{Inverting positive}\\\text{quantities reverses} \\\text{inequalities} \end{array}
\end{array}
\]
Therefore $b_n\geq a_n$. In Problem \ref{problemConvergencesum_2^infty1/(xlnx)dx} we illustrated (using the integral test) that $ \sum\limits_{n=2}^\infty (3b_n) $ is divergent and therefore so is its constant multiple $\sum\limits_{n=2}^\infty b_n$. Therefore $ \sum \limits_{n=2 }^\infty \frac{1}{(2n+1)\ln n}$ is divergent by the comparison test. 
}