\documentclass{article}
\usepackage{amsmath, amsfonts, amssymb, verbatim, hyperref}
\usepackage{enumitem}
\usepackage{pst-plot}
\usepackage{pstricks}

%\addtolength{\hoffset}{-3.5cm}
%\addtolength{\textwidth}{6.8cm}
%\addtolength{\voffset}{-3.3cm}
%\addtolength{\textheight}{6.3cm}

\usepackage{pdfpages}

\title{Math 240/242 Multivariable/Vector Calculus \\ Fall 2014}
\date{}
\begin{document}

%\color{green}
\maketitle
%\noindent\textbf{Time and place.}
%Monday, Wednesday, Friday 10-10:50, McCormack, Room 417, first floor. Monday 11:00-11:50,

\noindent \textbf{Instructor.} Todor Milev, \href{mailto:todor.milev@umb.edu}{\nolinkurl{todor.milev@umb.edu}} \quad \quad \quad .

\medskip
\noindent \textbf{Office hours. } Office hours by appointment Monday, Wednesday, Friday. Walk in office hours, Monday, Wednesday, Friday from 14:20-17:00 - only if I am in the office. Room: S-03-65.

\medskip
\noindent \textbf{Online resources. }  \url{https://piazza.com/umb/fall2014/math242240/home}  \quad \quad \quad .


\medskip\noindent \textbf{Textbook. }  James Stewart, Multivariable Calculus, Concepts and Contexts, ISBN 0-534-41004-9.

\medskip \noindent \textbf{Lecture slides. }  \url{https://piazza.com/umb/fall2014/math242240/home/resources} \quad \quad \quad .

\medskip\noindent Lecture slides will become available as the course progresses.

%\medskip
%\noindent \textbf{Prerequisite. } A standard pre-calculus course or equivalent.


\medskip
\noindent \textbf{Grades.} Your grade will consist of two tests, a comprehensive final exam, and a number of quizzes. 
\begin{itemize}
\item The quizzes will account for 20\% of your total grade.
\item The tests will account for 50\% (25\% each) of your total grade.
\item The final will account for 30\% of your total grade.
\end{itemize}
Please note that missed tests can not be made up, unless there is a valid medical reason accompanied with an official signed document from a medical doctor. Letter grades will be assigned as follows. 

\begin{center}
\begin{tabular}{lc|lc}
A & 85-100 & C & 65-69 \\
A-& 82-84 & C- & 62-64 \\
B+& 80-81 & D+ & 60-61 \\
B & 75-79& D & 55-59\\
B-& 72-74& D- & 50-54\\
C+& 70-71& F & below 50\\
\end{tabular}

\end{center}

No books, notes, calculators or any other electronic device (such as mobile phones) are allowed during any exam unless otherwise stated.

\medskip
\noindent \textbf{Homework.} You will be assigned homework, which will be posted on

\url{https://piazza.com/umb/fall2014/math141/resources} \quad \quad \quad .

\noindent I will expect you to complete the homework in written form in a convenient for you format (notebook, folder, etc.). However, \textbf{I will not check/collect/proofread homework.} 
 
\medskip
\noindent \textbf{Quizzes.} You will be given quizzes in class. The quiz times will be announced in class or on piazza.com. Quizzes may be announced from one lecture day to the next. Your quiz problem will be one of your homework problems, verbatim. 

\medskip
\noindent \textbf{Student conduct.} Students  are required to adhere the University Policy on Academic Standards and Cheating, to the University Statement of Plagiarism and the Documentation of Written Work, and to the Code of Student Conduct as described in the catalog of Undergraduate programs, pages 44-45 and 48-52. The code is available at the following web-page.

\noindent\url{http://www.umb.edu/life_on_campus/policies/code/}

\medskip
\noindent \textbf{Topics to be covered.} 
\renewcommand{\theenumii}{\arabic{enumii}}
\begin{enumerate}
\item Space, coordinates and coordinate systems.
\begin{enumerate}
\item Polar coordinates. 
\item Cylindrical coordinates.
\item Spherical coordinates.
\end{enumerate}
\item Vectors.
\begin{enumerate}
\item Scalar (dot) product.
\item Projections.
\item Cross product.
\item Scalar triple product.
\item Space orientation.
\end{enumerate}
\item Lines, Planes.
\item Functions of several variables.
\begin{enumerate}
\item Descriptions (graphical, numerical, analytical).
\item Slices and level curves.
\item Vector fields.
\item Quadratic surfaces.
\item Implicit functions.
\item Parametrizations.
\end{enumerate}
\item Parametrized curves.
\begin{enumerate}
\item Parametric curve equations.
\item Tangents.
\item Curve integrals.
\item Curve length.
\item Parametrization by arc length.
\end{enumerate}
\item Limits of functions of several variables.
\item Differentiation of functions of several variables.
\begin{enumerate}
\item Directional derivatives.
\item Partial derivatives.
\item Tangent plane.
\item Differentiability.
\item Differentials (total).
\item Multivariable chain rule.
\item Critical points, minima and maxima.
\end{enumerate}
\item Double integrals.
\item Triple integrals.
\item Iterated integrals.
\item Fubini's Theorem.
\item Curvilinear integrals.
\item Parametrized surfaces.
\item Surface integrals.
\end{enumerate}
\end{document}