\begin{frame}
\frametitle{A working system setup}
\begin{itemize}
\item To make presentation with freecalc/beamer/\LaTeX, you need to set-up a working environment.
\item Our system works on Windows, Mac and Linux. 
\item The system on which this presentation was prepared had the following software.
\begin{itemize}
\item Operational System (OS): Ubuntu Linux.
\item \LaTeX{} distribution: texlive.
\item \LaTeX{} editor: TeXstudio.
\item Subversion.
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Installing texlive}

\begin{itemize}
\item To install TeXlive on Ubuntu Linux, open terminal and type the command:
\begin{verbatim}
sudo apt-get install texlive-latex-extra
\end{verbatim}

\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Installing TeXstudio}

\begin{itemize}
\item To install TeXstudio, google ``texstudio''.
\item Navigate to the project's website.
\item Follow the download instructions for your system.
\end{itemize}
\end{frame}
\begin{frame}[fragile]
\frametitle{Installing Subversion}

\begin{itemize}
\item Subversion is a tool needed to fetch all freecalc files.
\item To install subversion open a terminal and type the command:
\begin{verbatim}
sudo apt-get install subversion
\end{verbatim}
\end{itemize}
\end{frame}
\begin{frame}[fragile]
\frametitle{Fetch the freecalc project files}
\begin{itemize}
\item You can fetch the freecalc project files from the Internet.
\item To do so, in your terminal, navigate to the folder where you want the files to be located. Then type the command:
\begin{verbatim}
svn co https://github.com/tmilev/freecalc.git
\end{verbatim}
\item In the not-so-likely occasion that the location of the freecalc has moved, in the above command you need to replace the URL \url{https://github.com/tmilev/freecalc.git} by the new updated location.
\item To find the freecalc project, do the following.
\begin{itemize}
\item Navigate to \url{github.com}.
\item Use the search box to search for ``freecalc''.
\item Of all projects named freecalc, select the one that belongs to Todor Milev (username tmilev).
\item Click on the ``Clone or download'' button to get the desired address.
\end{itemize}
\end{itemize}
\end{frame}