\documentclass{article}
\usepackage{../homework-problems-UMB}

\toggletrue{solutions}
%\togglefalse{solutions}
\toggletrue{answers}

\begin{document}
\begin{center}
\Large
Review problems Final\\ Summer 2014, Math 141 Calculus II \\
\end{center}

%\noindent \textbf{Name:} \hfill{~}
%\begin{tabular}{c|c|c|c|c|c|c|c|c||c}
%Problem&1 &2&3&4&5&6&7&8& $\sum$\\ \hline
%Score&&&&&&&&&\\ \hline
%Max&20&20&20&20&20&10&20&20&150
%\end{tabular} 

\noindent No calculators will be allowed. You will be expected to finish in 110 minutes. The time for work will be up to 3 hours. We will move rooms with the students who do not finish the test in the first 110 minutes.

For the final, you are allowed to bring up to 4 double sided pages, hand-written by you. The pages are allowed to contain any information you think is relevant to the test, including entire problem solutions. You may write complete solutions of the review problems on the sheets.

The 4 pages will be collected with the test. If the pages are more than 4 or they are not hand-written by you, this will be considered cheating and you will get 0 on the final. 
\section{Practice problems}

\begin{enumerate}
\item \input{../../modules/lhospital/homework/lhospital-rule-final-review-collection-1}
\item \input{../../modules/lhospital/homework/lhospital-rule-coming-from-Maclaurin-series-2}
\item \input{../../modules/partial-fractions/homework/integration-rational-functions-complete-3}
%\item \input{../../modules/partial-fractions/homework/integration-rational-functions-complete-4}
\item \input{../../modules/trig-substitution/homework/integration-euler-substitution-case1}

%\item \input{../../modules/integration/homework/integral-problem-collection-all-techniques-1}

\item \input{../../modules/improper-integrals/homework/improper-integral-convergent-or-divergent-3}
\item \input{../../modules/improper-integrals/homework/improper-integral-convergent-or-divergent-4}


%\item \input{../../modules/parametric-curves/homework/parametric-curves-find-derivatives-1}

%\item \input{../../modules/parametric-curves/homework/curve-length-2}

%\item \input{../../modules/polar-curves/homework/area-locked-by-curve-2}

%\item \input{../../modules/polar-curves/homework/area-locked-by-curve-1}

\item \input{../../modules/series/homework/geometric-series-sum-1}

%\item  \input{../../modules/series/homework/series-convergent-or-divergent-all-techniques-1}

%\item \input{../../modules/power-series/homework/interval-of-convergence-1}

%\item \input{../../modules/power-series/homework/find-taylor-series-1}

\item 
\input{../../modules/power-series/homework/find-maclaurin-series-1}

%\item \input{../../modules/power-series/homework/find-maclaurin-series-2}

%\item \input{../../modules/power-series/homework/find-maclaurin-series-3}

\item 
\input{../../modules/power-series/homework/find-maclaurin-series-4}

\item 
\input{../../modules/power-series/homework/find-maclaurin-series-5}

%\item \input{../../modules/power-series/homework/find-maclaurin-series-6}

\item \input{../../modules/diff-eq-separable/homework/separable-DFQs-3}

%\item  \input{../../modules/complex-numbers/homework/polar-form-complex-numbers-1}

%\item  \input{../../modules/complex-numbers/homework/find-nth-root-1}
\end{enumerate}

\section{Solutions to some of the problems}
\input{../../modules/partial-fractions/homework/integration-rational-functions-complete-3-solutions}


\input{../../modules/trig-substitution/homework/integration-euler-substitution-case1-solutions}

\input{../../modules/series/homework/geometric-series-sum-1-solution}



\end{document}





