\begin{frame}
\begin{itemize}
\item We present the aggregate signature scheme from \cite[MPSW18]{cryptoeprintSimpleSchnorrMultisignatures}.
\item Different from Zilliqa in one crypto step (makes it more secure).
\item The scheme does not specify communication protocol.
\item For signing, we propose to combine \cite[MPSW18]{cryptoeprintSimpleSchnorrMultisignatures} with the signing protocol of Zilliqa.
\item For initial setup and general outline of the networking we present our own setup that leverages the foundation chain.



\end{itemize}
\end{frame}

\begin{frame}
Secr.: $\mathrm{secret}_1, \dots, \mathrm{secret}_n$, pub. keys: $\mathrm{pub}_1 = g^{ \mathrm{secret }_1 }, \dots, \mathrm{pub}_n= g^{\mathrm{secret}_n}$.
\begin{emptyTheorem}[{\cite[MPSW18]{cryptoeprintSimpleSchnorrMultisignatures} aggregate signature}: signing]
\begin{itemize}
\item Aggregator: send protocol start to signers.
\item Each signer: Choose random $\mathrm{nonce}_i$, compute $q_i = g^{\mathrm{nonce}_i}$.
\item Each signer: send $q_i$ to aggregator. Let $\mathcal A$: set of healthy nodes.
\item Aggregator: compute $\alertNoH{2}{a_i= H(\mathcal A, \mathrm{pub}_i)}$. Compute $ \mathrm{Pub} =\prod_{i\in \mathcal A} \mathrm{pub}_i^{\alertNoH{2}{a_i}}$. 
\item Aggregator: compute $Q = \prod_{i \in \mathcal A} q_i$.
\item Aggregator: compute  $\mathrm{challenge} = H(Q, \mathrm{Pub}, \mathrm{digest})$.
\item Aggregator: send $\mathrm{challenge},\mathrm{Pub}, \mathrm{digest}$ to signers. Bad net: reset.
\item Each signer: verify $\mathrm{challenge} = H(Q, \mathrm {Pub}, \mathrm{digest}) $. Bad: reset.
\item Each signer: compute $\mathrm{solution}_i = {\mathrm{nonce}_i - \mathrm{challenge} \cdot \mathrm{secret}_i} $.
\item Each signer: send $\mathrm{solution}_i$ to aggregator. Bad net: reset.
\item Aggregator: compute $\mathrm{solution} = \sum_i \mathrm{solution}_i $.
\item Aggregator: final signature: $(\mathrm{challenge}, \mathrm{solution}), \mathcal A$.
\end{itemize}
\end{emptyTheorem}
\vskip -0.15cm
Like Zilliqa except $a_i$'s: those give extra security (Wagner-alg. attack).

\vskip 15cm

\end{frame}

\begin{frame}
\begin{itemize}
\item Signature verification for \cite{cryptoeprintSimpleSchnorrMultisignatures} is similar to Zilliqa's and we omit it.
\end{itemize}
\end{frame}

\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
\node[shape=circle,draw,inner sep=1pt] (char) {\tiny #1};}}

\begin{frame}
\begin{itemize}
\item Intended use case: aggregate signature for each transaction, completed within 0-5 seconds after initiation. Sample networking:

\hfil\hfil$
\xymatrix{ 
& &
\textrm{kb}_1
&\textrm{kb}_2 \\
&\mathrm{kb}_8 &&&\textrm{kb}_3\\
\mathrm{wallet} 
\ar@/^0.5pc/@{->}[r]^{\circled{1}}  
\ar@/_0.5pc/@{<-}[r]_{\circled{6}}  &
\mathrm{kb}_7 
\ar@/^0.5pc/@{=>}[u]_{\dots} 
\ar@/_0.5pc/@{<=}[u] 
\ar@/^0.5pc/@{=>}[uur]_{\dots}%^{\circled{2}, \circled{4}}_{\circled{3}, \circled{5}} 
\ar@/_0.5pc/@{<=}[uur]
\ar@/^0.5pc/@{=>}[uurr]_{\dots}%^{\circled{2}, \circled{4}}_{\circled{3}, \circled{5}} 
\ar@/_0.5pc/@{<=}[uurr]_{\dots}%^{\circled{2}, \circled{4}}_{\circled{3}, \circled{5}} 
\ar@/^0.5pc/@{=>}[rrru]_{\dots}
\ar@/_0.5pc/@{<=}[rrru]
\ar@/^0.5pc/@{=>}[rrr]^(.65){\circled{2}, \circled{4}} 
\ar@/_0.5pc/@{<=}[rrr]_(.65){\circled{3}, \circled{5}}
\ar@/^0.5pc/@{=>}[rrd]_{\dots}%^{\circled{2}, \circled{4}}_{\circled{3}, \circled{5}} 
\ar@/_0.5pc/@{<=}[rrd]_{\dots}%^{\circled{2}, \circled{4}}_{\circled{3}, \circled{5}} 
\ar@/^0.5pc/@{=>}[rd]_{\dots}%^{\circled{2}, \circled{4}}_{\circled{3}, \circled{5}} 
\ar@/_0.5pc/@{<=}[rd]_{\dots}%^{\circled{2}, \circled{4}}_{\circled{3}, \circled{5}} 
&&&\textrm{kb}_4\\
& &\textrm{kb}_6&\textrm{kb}_5 \\
}
$
\begin{itemize}
\item \circled{1} Wallet sends transaction ``$A$ sends $B$ amount, signed by $A$''. Node $\mathrm{kb}_1$ is designates itself as an aggregator.
\item \circled{2} Aggregator starts protocol, requesting the $q_i$'s
\item \circled{3} Signers report their $q_i$'s, proving they are online.
\item \circled{4} Aggregator sends challenges out.
\item \circled{5} Signers reply with solutions to their challenges.
\item \circled{6} Aggregator reports signature to wallet.
\end{itemize}
\end{itemize}
\end{frame}
