\documentclass{article}
\usepackage{amsmath, amsfonts, amssymb, verbatim, hyperref}
\usepackage{enumitem}
\usepackage{pst-plot}
\usepackage{pstricks}

%\addtolength{\hoffset}{-3.5cm}
%\addtolength{\textwidth}{6.8cm}
%\addtolength{\voffset}{-3.3cm}
%\addtolength{\textheight}{6.3cm}

\newcommand{\websitebase}{https://piazza.com/umb/summer2015/math140}

\usepackage{pdfpages}

\title{Math 141 Calculus II \\ Summer 2015}
\date{}
\begin{document}

%\color{green}
\maketitle
%\noindent\textbf{Time and place.}
%Monday, Wednesday, Friday 10-10:50, McCormack, Room 417, first floor. Monday 11:00-11:50,

\noindent \textbf{Instructor.} Todor Milev, \href{mailto:todor.milev@umb.edu}{\nolinkurl{todor.milev@umb.edu}} \quad \quad \quad .

\medskip
\noindent \textbf{Office hours. } Office hours by appointment Tuesday, Thursday. Walk in office hours, Tuesday, Thursday - 14:30-16:30 only if I am in the office. Room: S-03-65.

\medskip
\noindent \textbf{Online resources. }  \url{\websitebase/home}  \quad \quad \quad .


\medskip\noindent \textbf{Textbook. }  James Stewart, Calculus, 7th edition, published by Brooks Cole, 2012. ISBN-13: 978-0-538-49781-7
ISBN-10: 0-538-49781-5.

\medskip \noindent \textbf{Lecture slides. }  \url{\websitebase/resources} \quad \quad \quad .

\medskip \noindent \textbf{Master problem sheet (latest version). }  \url{\websitebase/resources} \quad \quad \quad .


\medskip\noindent Lecture slides will become available as the course progresses.

%\medskip
%\noindent \textbf{Prerequisite. } A standard pre-calculus course or equivalent.


\medskip
\noindent \textbf{Grades.} Your grade will consist of two tests, a comprehensive final exam, and a number of quizzes. 
\begin{itemize}
\item The quizzes will account for 20\% of your total grade.
\item The tests will account for 50\% (25\% each) of your total grade.
\item The final will account for 30\% of your total grade.
\end{itemize}
Please note that missed tests can not be made up, unless there is a valid medical reason accompanied with an official signed document from a medical doctor. Letter grades will be assigned as follows. 

\begin{center}
\begin{tabular}{lc|lc}
A & 85-100 & C & 65-69 \\
A-& 82-84 & C- & 62-64 \\
B+& 80-81 & D+ & 60-61 \\
B & 75-79& D & 55-59\\
B-& 72-74& D- & 50-54\\
C+& 70-71& F & below 50\\
\end{tabular}

\end{center}

No books, notes, calculators or any other electronic device (such as mobile phones) are allowed during any exam unless otherwise stated.

\medskip
\noindent \textbf{Homework.} You will be assigned homework, which will be posted on

\url{https://piazza.com/umb/spring2015/m141/resources} \quad \quad \quad .

\noindent I will expect you to complete the homework in written form in a convenient for you format (notebook, folder, etc.). However, \textbf{I will not check/collect/proofread homework.} 
 
\medskip
\noindent \textbf{Quizzes.} You will be given quizzes in class. \textbf{The time of the quiz will be announced in class}. Quizzes may be announced from one lecture day to the next. Your quiz problem will be one of your homework problems, verbatim. 

\medskip
\noindent \textbf{Student conduct.} Students  are required to adhere the University Policy on Academic Standards and Cheating, to the University Statement of Plagiarism and the Documentation of Written Work, and to the Code of Student Conduct as described in the catalog of Undergraduate programs, pages 44-45 and 48-52. The code is available at the following web-page.

\noindent\url{http://www.umb.edu/life_on_campus/policies/code/}

\medskip
\noindent \textbf{Topics to be covered.} 

\medskip
\noindent \textbf{List of topics from previous years.} The list of topics is a preliminary guideline, and will be subject to change.
\begin{enumerate}[label*=\arabic*.]
\item Functions and limits
\begin{enumerate}[label*=\arabic*.]
\item Ways to represent a function.
\item Some essential functions.
\item New functions from old.
\item Tangents; velocity problems.
\item Limit of a function.
\item Calculating limits with limit laws.
\item Continuity.
\end{enumerate}
\item Derivatives
\begin{enumerate}[label*=\arabic*.]
\item Derivatives and rates of change.
\item The derivative as a function.
\item Differentiation formulas.
\item Derivatives of trigonometric functions.
\item The chain rule.
\item Implicit differentiation.
\item Some applications of rates of change.
\item Related rates.
\item Linear approximations and differentials.
\end{enumerate}
\item Applications of differentiation.
\begin{enumerate}[label*=\arabic*.]
\item Maxima and minima.
\item The Mean Value Theorem.
\item Derivatives and shape of a graph.
\item Limits at $\infty$; asymptotes.
\item On curve sketching.
\item On optimization problems.
\item Newton's method.
\item Antiderivatives.
\end{enumerate}
\item Integrals
\begin{enumerate}[label*=\arabic*.]
\item Areas and distances.
\item Definite integrals.
\item The Fundamental Theorem of Calculus.
\item Indefinite integrals and the Net Change Theorem.
\item The substitution rule.
\end{enumerate}
\item First applications of integration.
\begin{enumerate}[label*=\arabic*.]
\item Area between curves.
\item Volumes of solids of revolution.
\end{enumerate}
\item Inverse functions.
\begin{enumerate}[label*=\arabic*.]
\item Inverse functions.
\item Exponential functions and their derivatives.
\item Logarithmic functions.
\item Derivatives of inverse functions.
\end{enumerate}
\end{enumerate}


\end{document}