\documentclass{article}
\usepackage{amsmath, amsfonts, amssymb, verbatim, hyperref}
\usepackage{enumitem}
\usepackage{pst-plot}
\usepackage{pstricks}

%\addtolength{\hoffset}{-3.5cm}
%\addtolength{\textwidth}{6.8cm}
%\addtolength{\voffset}{-3.3cm}
%\addtolength{\textheight}{6.3cm}

\usepackage{pdfpages}

\title{Math 141 Calculus II \\ Spring 2015}
\date{}
\begin{document}

%\color{green}
\maketitle
%\noindent\textbf{Time and place.}
%Monday, Wednesday, Friday 10-10:50, McCormack, Room 417, first floor. Monday 11:00-11:50,

\noindent \textbf{Instructor.} Todor Milev, \href{mailto:todor.milev@umb.edu}{\nolinkurl{todor.milev@umb.edu}} \quad \quad \quad .

\medskip
\noindent \textbf{Office hours. } Office hours by appointment Tuesday, Thursday. Walk in office hours, Tuesday, Thursday from 15:00-17:00 - only if I am in the office. Room: S-03-65.

\medskip
\noindent \textbf{Online resources. }  \url{https://piazza.com/umb/spring2015/m141/home}  \quad \quad \quad .


\medskip\noindent \textbf{Textbook. }  James Stewart, Calculus, 7th edition, published by Brooks Cole, 2012. ISBN-13: 978-0-538-49781-7
ISBN-10: 0-538-49781-5.

\medskip \noindent \textbf{Lecture slides. }  \url{https://piazza.com/umb/spring2015/m141/resources} \quad \quad \quad .

\medskip \noindent \textbf{Master problem sheet (latest version). }  \url{https://piazza.com/umb/spring2015/m141/resources} \quad \quad \quad .


\medskip\noindent Lecture slides will become available as the course progresses.

%\medskip
%\noindent \textbf{Prerequisite. } A standard pre-calculus course or equivalent.


\medskip
\noindent \textbf{Grades.} Your grade will consist of two tests, a comprehensive final exam, and a number of quizzes. 
\begin{itemize}
\item The quizzes will account for 20\% of your total grade.
\item The tests will account for 50\% (25\% each) of your total grade.
\item The final will account for 30\% of your total grade.
\end{itemize}
Please note that missed tests can not be made up, unless there is a valid medical reason accompanied with an official signed document from a medical doctor. Letter grades will be assigned as follows. 

\begin{center}
\begin{tabular}{lc|lc}
A & 85-100 & C & 65-69 \\
A-& 82-84 & C- & 62-64 \\
B+& 80-81 & D+ & 60-61 \\
B & 75-79& D & 55-59\\
B-& 72-74& D- & 50-54\\
C+& 70-71& F & below 50\\
\end{tabular}

\end{center}

No books, notes, calculators or any other electronic device (such as mobile phones) are allowed during any exam unless otherwise stated.

\medskip
\noindent \textbf{Homework.} You will be assigned homework, which will be posted on

\url{https://piazza.com/umb/spring2015/m141/resources} \quad \quad \quad .

\noindent I will expect you to complete the homework in written form in a convenient for you format (notebook, folder, etc.). However, \textbf{I will not check/collect/proofread homework.} 
 
\medskip
\noindent \textbf{Quizzes.} You will be given quizzes in class. The quiz times will be announced in class or on piazza.com. Quizzes may be announced from one lecture day to the next. Your quiz problem will be one of your homework problems, verbatim. 

\medskip
\noindent \textbf{Student conduct.} Students  are required to adhere the University Policy on Academic Standards and Cheating, to the University Statement of Plagiarism and the Documentation of Written Work, and to the Code of Student Conduct as described in the catalog of Undergraduate programs, pages 44-45 and 48-52. The code is available at the following web-page.

\noindent\url{http://www.umb.edu/life_on_campus/policies/code/}

\medskip
\noindent \textbf{Topics to be covered.} 
\renewcommand{\theenumii}{\arabic{enumii}}
\begin{enumerate}
\item Review of trigonometric functions.
\item Inverse trigonometric functions.
\begin{enumerate}
\item Definitions of inverse trigonometric functions.
\item Derivatives of inverse trigonometric functions.
\end{enumerate}
\item Integration, Review.
\begin{enumerate}
\item The Fundamental Theorem of Calculus.
\item Differential forms.
\item Integration and logarithms.
\end{enumerate}
\item Techniques of integration
\begin{enumerate}
\item Integration by parts.
\item Integration of rational functions.
\begin{enumerate}
\item Building block integrals.
\item Partial fractions.
\end{enumerate}
\item Trigonometric integrals.
\item Integrals of radicals of quadratics
\begin{enumerate}
\item Trigonometric substitutions.
\item Euler substitutions corresponding to trig substitutions.
\end{enumerate}
\end{enumerate}
\item L'Hospital's rule.
\item Improper integrals.
\item Sequences.
\item Series.
\begin{enumerate}
\item Geometric and arithmetic sums.
\item Telescoping series.
\item Comparison test.
\item Integral test.
\item Absolute convergence, alternating series.
\item Ratio and root tests.
\end{enumerate}
\item Power series.
\begin{enumerate}
\item Radius and interval of convergence.
\item Maclaurin and Taylor series.
\item Integrating and differentiating Maclaurin and Taylor series. 
\item Maclaurin series of $\ln(1+x), e^x, \sin x, \cos x, \arctan x, \arcsin x$.
\end{enumerate}
\item Polar coordinates.
\item Complex numbers (if time allows).
\item Curves
\begin{enumerate}
\item Curve images and parametric curves definitions.
\item Curve (arc) length.
\item Curves in polar coordinates.
\item Area locked by curves.
\end{enumerate}
\item Differential equations
\begin{enumerate}
\item Direction fields.
\item Separable equations.
\item The logistic (population growth) equation.
\end{enumerate}
\item Solids of revolution (if time allows).
\begin{enumerate}
\item Washer method (the volume form $\operatorname{d}\rho \operatorname{d}\theta \operatorname{d} z$, cylindrical coordinates).
\item Cylindrical shells method (the volume form $ \operatorname{d}\theta \operatorname{d} z \operatorname{d}\rho$, cylindrical coordinates).
\end{enumerate}
\end{enumerate}
\end{document}