\documentclass{article}
\usepackage{../homework-problems-UMB}

\toggletrue{solutions}
%\togglefalse{solutions}
\toggletrue{answers}

\begin{document}
\begin{center}
\Large
Review problems Test 1\\ Math 240/242 Calculus III \\ \normalsize Instructor: Todor Milev
\end{center}

%\noindent \textbf{Name:} \hfill{~}
%\begin{tabular}{c|c|c|c|c|c|c|c|c||c}
%Problem&1 &2&3&4&5&6&7&8& $\sum$\\ \hline
%Score&&&&&&&&&\\ \hline
%Max&20&20&20&20&20&10&20&20&150
%\end{tabular} 

\noindent The exam is closed books, no calculators will be allowed. You are allowed one single-sided formula sheet, handwritten by you. The sheet will be collected with the test. \textbf{The sheet can contain only formulas. Solved sample problems are not acceptable.}


Except possibly one problem, the problem types on the exam will be selected from problem types covered in the homeworks so far. This includes the homeworks which were not quizzed. \textbf{I may give you one theoretical question or problem not of type covered in the review sheet. To be prepared to answer such a question, you need to study the slides and/or the textbook.} Your review sheet contains only a few new problems. It mostly consists of an aggregate of homework problems you have been previously assigned. 

\begin{enumerate}
\item 
\input{../../modules/partial-derivatives/homework/compute-tangent-plane-graph-function-1}
\item  \input{../../modules/partial-derivatives/homework/compute-tangent-plane-parametric-surface-1}
\item \input{../../modules/partial-derivatives/homework/compute-partial-derivative-1}

\input{../../modules/partial-derivatives/homework/compute-partial-derivative-1-solutions}

\item \input{../../modules/optimization-multivariable/homework/find-min-max-1}
\item \input{../../modules/optimization-multivariable/homework/min-max-with-lagrange-multipliers}

\item \input{../../modules/integration-multivariable/homework/double-integrals-1}

\end{enumerate}

\end{document}