\documentclass{article}
\input{../homework-problems-UMB}

\toggletrue{solutions}
%\togglefalse{solutions}
\toggletrue{answers}

\newtheorem{problem}{Problem}
\begin{document}
\begin{center}
\Large
Review Problems, Final Exam\\ Math 240/242 Calculus III \\ \normalsize Instructor: Todor Milev
\end{center}

%\noindent \textbf{Name:} \hfill{~}
%\begin{tabular}{c|c|c|c|c|c|c|c|c||c}
%Problem&1 &2&3&4&5&6&7&8& $\sum$\\ \hline
%Score&&&&&&&&&\\ \hline
%Max&20&20&20&20&20&10&20&20&150
%\end{tabular}

\noindent The exam is closed books, no calculators will be allowed. You are allowed one single-sided formula sheet, handwritten by you. The sheet will be collected with the test. \textbf{The sheet can contain only formulas. Solved sample problems are not acceptable.}

Your review sheet consists mainly of problems that were included in the lecture slides. Except possibly one problem, the problem types on the exam will be selected from problem types covered in this review sheet. 


\textbf{I may give you one theoretical question or problem not of type covered in the review sheet. To be prepared to answer such a question, you need to study the slides and/or the textbook.} 
\section{Lines, Planes, Distance}
\begin{problem}
\input{../../modules/vectors/homework/distance-between-lines}
\end{problem}
\input{../../modules/vectors/homework/distance-between-lines-solutions}


\section{Tangent Planes}
\begin{problem}
\input{../../modules/partial-derivatives/homework/compute-tangent-plane-surface-implicit-equation-1}
\end{problem}
\input{../../modules/partial-derivatives/homework/compute-tangent-plane-surface-implicit-equation-1-solutions}

\section{Minima Maxima}
\begin{problem}
\input{../../modules/optimization-multivariable/homework/find-min-max-1}
\end{problem}

\section{Triple Integrals}
\begin{problem}
\input{../../modules/integration-multivariable/homework/triple-integrals-over-curvilinear-trapezoids-1}
\end{problem}
\input{../../modules/integration-multivariable/homework/triple-integrals-over-curvilinear-trapezoids-1-solutions}

\section{Variable Changes in Multivariable Integrals}
\begin{problem}
\textbf{Problem \ref{problemVolumeRaisedHorn} is of higher difficulty than the problem you will get on the exam.}

\input{../../modules/integration-multivariable-variable-change/homework/volumes-1}
\end{problem}

\subsection{Integrals in Polar, Cylindrical and Spherical coordinates}
\subsubsection{Polar coordinates}
\begin{problem}
\input{../../modules/integration-multivariable-variable-change/homework/integrals-polar-1}
\end{problem}
\subsubsection{Cylindrical coordinates}
\begin{problem}
\input{../../modules/integration-multivariable-variable-change/homework/integrals-cylindrical-1}
\end{problem}
\subsubsection{Spherical coordinates}
\begin{problem}
\input{../../modules/integration-multivariable-variable-change/homework/integrals-spherical-centroids-centers-of-mass-1}
\end{problem}
\section{2D Field Potential}
\begin{problem}
\input{../../modules/integration-line-integrals/homework/find-2D-scalar-potential}
\end{problem}
\section{Green's Theorem}
\subsection{Line integrals via Greens Theorem}
\begin{problem}
\input{../../modules/integration-line-integrals/homework/line-integrals-via-greens-theorem}
\end{problem}
\section{Surface Integrals}
\subsection{Surface area}
\begin{problem}
\input{../../modules/integration-surface-integrals/homework/surface-area-1}
\end{problem}
\subsection{Flux}
\begin{problem}
\input{../../modules/integration-surface-integrals/homework/find-flux-1}
\end{problem}
\subsection{Divergence Theorem }
\begin{problem}
\input{../../modules/integration-surface-integrals/homework/divergence-theorem-1}
\end{problem}

\end{document}
