\documentclass{article}
\input{../homework-problems-UMB}

\toggletrue{solutions}
\toggletrue{answers}
\renewcommand{\fcProblemRef}{\theproblem.\theenumi}
\renewcommand{\fcSubProblemRef}{\theproblem.\theenumi.\theenumii}


\newcommand{\hide}[1]{}
\newtheorem{problem}{Problem}
\pagestyle{empty}
\begin{document}
\begin{center}
\Large
Review sheet Test 2 \\ Math 141 Calculus II \\ \normalsize Spring 2016 \\ Instructor(s): Todor Milev
\end{center}
%\noindent \textbf{Name:\underline{~~~~~~~~~~~~~~~~~~~~~~~} } \hfill{~}



\noindent The exam is closed textbook. \textbf{No electronic devices are allowed during the exam. } The exam will contain a number of problems and sub-problems (the number of problems may depend on your section/instructor). \textbf{Expect each problem type from this review sheet to appear on the test.}%You are allowed one single formula sheet, handwritten by you. No template problem solutions are allowed. The sheet will be collected with the test. Photocopied formula sheets are not allowed. 

\begin{problem}Compute the limit.
\begin{enumerate}[ref={\fcProblemRef}]
\item \input{../../modules/lhospital/homework/lhospital-rule-coming-from-Maclaurin-series-1-problem-12}
\item \input{../../modules/lhospital/homework/lhospital-rule-coming-from-Maclaurin-series-1-problem-14}
\item \input{../../modules/lhospital/homework/lhospital-rule-coming-from-Maclaurin-series-1-problem-15}
\end{enumerate}
\end{problem}
\input{../../modules/lhospital/homework/lhospital-rule-coming-from-Maclaurin-series-1-solutions}

\begin{problem} Determine whether the integral is convergent or divergent. If convergent, evaluate it. Motivate your answer.

\begin{multicols}{2}
\begin{enumerate}[ref={\fcProblemRef}]
\item \input{../../modules/improper-integrals/homework/improper-integral-convergent-or-divergent-1-problem-8}
\item \input{../../modules/improper-integrals/homework/improper-integral-convergent-or-divergent-1-problem-12}
\item \input{../../modules/improper-integrals/homework/improper-integral-convergent-or-divergent-1-problem-13}
\item \input{../../modules/improper-integrals/homework/improper-integral-convergent-or-divergent-1-problem-19}
\end{enumerate}
\end{multicols}
\end{problem}
\input{../../modules/improper-integrals/homework/improper-integral-convergent-or-divergent-1-solutions}
\begin{problem}
\begin{enumerate}[ref={\fcProblemRef}]
\item \input{../../modules/parametric-curves/homework/curve-length-1-problem-2}
\item \input{../../modules/parametric-curves/homework/curve-length-1-problem-3}
\item \input{../../modules/parametric-curves/homework/curve-length-1-problem-6}
\item \input{../../modules/parametric-curves/homework/curve-length-1-problem-7}
\end{enumerate}
\end{problem}
\input{../../modules/parametric-curves/homework/curve-length-1-solutions}


\begin{problem}
\input{../../modules/polar-curves/homework/area-locked-by-curve-1}

\end{problem}
\input{../../modules/polar-curves/homework/area-locked-by-curve-1-solutions}



%\begin{problem}
%\input{../../modules/sequences/homework/convergent-or-divergent-sequence-1}
%\end{problem}
%\input{../../modules/sequences/homework/convergent-or-divergent-sequence-1-solutions}


%\begin{problem}
% \input{../../modules/series/homework/geometric-series-sum-1} 
%\end{problem}
%\input{../../modules/series/homework/geometric-series-sum-1-solutions}
%\begin{problem}
%\input{../../modules/series/homework/series-basic-tests-1}
%\end{problem}
%\input{../../modules/series/homework/series-basic-tests-1-solutions}

%\begin{problem}
%\input{../../modules/series/homework/series-integral-or-comparison-test-1}
%\end{problem}
%\input{../../modules/series/homework/series-integral-or-comparison-test-1-solutions}



%\vskip 18cm
%\hfill \begin{tabular}{c|c|c|c|c|c|c||c}
%Problem&1 &2&3&4&5&6& $\sum$\\ \hline
%Score&&&&&&&\\ \hline
%Max&17&17&17&17&17&17&102
%\end{tabular} 


\end{document}