\documentclass{article}
\input{../homework-problems-UMB}
\renewcommand{\fcProblemRef}{\theenumi.\alph*}

\toggletrue{solutions}
\toggletrue{answers}
%\togglefalse{solutions}

\renewcommand{\course}{Math 130}
\newcommand{\tooeasy}{ \textbf{The problem is too easy to appear on a quiz or test. }}

\begin{comment}
\homeworkStart{on Lecture 1 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/coordinate-systems/homework/plot-the-points-2d-1}
\item \input{../../modules/coordinate-systems/homework/distance-between-points-2d-1}
\input{../../modules/coordinate-systems/homework/distance-between-points-2d-1-solutions}
\item \input{../../modules/coordinate-systems/homework/find-midpoint-2d-1}

\homeworkEnd
\end{comment}

\begin{comment}
\homeworkStart{on Lecture 2 \\ Quiz time will be announced in class.}{}

\item \input{../../modules/coordinate-systems/homework/recognizing-circles-1}
\input{../../modules/coordinate-systems/homework/recognizing-circles-1-solutions}
\item \input{../../modules/coordinate-systems/homework/equation-circle-from-center-and-point-1}

\item \input{../../modules/equation-graph/homework/find-x-y-intercepts-graph-given-circle-1}

\homeworkEnd
\end{comment}
\begin{comment}
\homeworkStart{on Lecture 3 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/lines-2d/homework/line-from-two-points-1}
\item \input{../../modules/lines-2d/homework/find-line-intersection-1}
\input{../../modules/lines-2d/homework/find-line-intersection-1-solutions}

\homeworkEnd
\end{comment}

\begin{comment}
\homeworkStart{on Lecture 4 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/precalculus/homework/functions-evaluate-simplify}
\item \input{../../modules/precalculus/homework/functions-domains-ranges-1}
\item \input{../../modules/functions-basics/homework/piecewise-linear-formula-from-plot-1}
\item \input{../../modules/functions-basics/homework/even-or-odd-from-formula-1}
\input{../../modules/functions-basics/homework/even-or-odd-from-formula-1-solutions}

%\item \input{../../modules/functions-basics/homework/piecewise-linear-piecewise-circular-formula-from-plot-1}


\homeworkEnd
\end{comment}

\begin{comment}
\homeworkStart{on Lecture 5 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/functions-basics/homework/functions-composing-fractional-linear-1}
\item \input{../../modules/functions-basics/homework/function-composition-domains-and-ranges-1}

\homeworkEnd
\end{comment}
\begin{comment}
\homeworkStart{on Lecture 6 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/function-graph-linear-transformations/homework/use-known-graph-to-sketch-its-linear-transformation-1}
 \input{../../modules/function-graph-linear-transformations/homework/use-known-graph-to-sketch-its-linear-transformation-1-solutions}
\homeworkEnd
\end{comment}

%\begin{comment}
\homeworkStart{on Lecture 7 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/quadratic-functions/homework/solve-the-quadratic-equation-1}
\item \input{../../modules/quadratic-functions/homework/plot-the-quadratic-1}
\item \input{../../modules/quadratic-functions/homework/maxima-minima-quadratic-1}
\homeworkEnd
\end{comment}

\begin{comment}
\homeworkStart{on Lecture 8 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/exponential-functions/homework/exponent-simplify}
\input{../../modules/exponential-functions/homework/exponent-simplify-solutions}
\item \input{../../modules/exponential-functions/homework/sketch-exponent-shift-from-known-exponent-1}
\item \input{../../modules/logarithms/homework/e-as-a-limit-and-compound-interest}
\input{../../modules/logarithms/homework/e-as-a-limit-and-compound-interest-solutions}

\homeworkEnd
\end{comment}

\begin{comment}
\homeworkStart{on Lecture 9 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/inverse-functions/homework/inverse-functions3}
\input{../../modules/inverse-functions/homework/inverse-functions3-solutions}
\item \input{../../modules/inverse-functions/homework/inverse-functions4}

\item \textbf{This problem uses material that has not been studied (yet), and will therefore not appear on the quiz.}
\input{../../modules/inverse-functions/homework/inverse-functions2}
\input{../../modules/inverse-functions/homework/inverse-functions2-solutions}

\homeworkEnd
\end{comment}
\begin{comment}
\homeworkStart{on Lecture 10 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/logarithms/homework/logarithms-basic-no-properties-1}
\homeworkEnd
\end{comment}

%\begin{comment}
\homeworkStart{on Lecture 11 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/logarithms/homework/logarithms-basic-properties-1}
\input{../../modules/logarithms/homework/logarithms-basic-properties-1-solutions}
\homeworkEnd
\end{comment}
