\documentclass{article}
\usepackage{amsmath, amsfonts, amssymb, verbatim, hyperref}
\usepackage{auto-pst-pdf}
\usepackage{pst-plot}
\addtolength{\hoffset}{-3.5cm}
\addtolength{\textwidth}{6.8cm}
\addtolength{\voffset}{-3.3cm}
\addtolength{\textheight}{6.3cm}
\renewcommand{\Re}{\mathrm{Re~}}
\renewcommand{\Im}{\mathrm{Im~}}
\newcommand{\doublebrace}[4]{\left\{\begin{array}{ll} #1 & #2 \\#3 & #4  \end{array} \right.}

\date{}
\newtheorem{problem}{Problem}
\title{Exam I\\ Math 140 Calculus I }
\pagestyle{empty}
\begin{document}
\begin{center}
\LARGE
Exam I 

Calculus I, Math 140

March 7, 2013
\end{center}

\noindent \textbf{Name:}
\medskip
%\maketitle
%\color{green}

\begin{problem}
(7.5 + 7.5 pts) Find an expression for the function $(f\circ g)(x)$ and $(g\circ f)(x)$, where 
\[
f(x)= \frac{3x-1}{x-2},\quad  g(y)=\frac{y-2 }{2y-4} 
\] Simplify your answer to a single fraction.
\end{problem}

\vskip 9cm

\begin{problem} (15 pts)
Find all solutions in the interval $[0,2\pi]$ of the equation.
$\sqrt{ 3}\sin x=  \sin 2x$
\end{problem}

\newpage
\begin{problem} (15 pts) Evaluate the limit if it exists.
\[\lim\limits_{x\to 2} \frac{3x^2-4x-4}{x^3-4x}
\]
\end{problem}

\vskip 9cm

\begin{problem} (10 pts)
Evaluate the limit.
\[\lim\limits_{x\to 2^+} \frac{\sqrt{x^3-4x }}{3x^2 -4x-4 }\quad .
\]
\end{problem}

\newpage
\begin{problem} (10 pts)
Use the intermediate value theorem to show that the equation has a solution in the interval $(-2,0)$.
\[e^{2x}+x+1 =\sin  (-x) \quad .
\]
\end{problem}
\vskip 9cm
\begin{problem} (10pts)
Solve the equation.
\[e^{-4x}+5e^{-2x}-6=0\quad .
\]
\end{problem}
\newpage
\begin{problem} (15 pts)
Plot roughly the function 
\[
f(x)= x^{2}+2x+2
\] with domain $x>-1$. Compute the inverse function $f^{-1}(y) $. Plot roughly $f^{-1}(x)$. Explain what is the relationship between the graphs of $f^{-1}(x)$ and $f(x)$.
\end{problem}
\vskip 9cm
\begin{problem} (10 pts)
Compute the limit.
\[\lim\limits_{x\to\infty}\sqrt{x^2+3x}-\sqrt{x^2-\frac{x}3}\quad .
\]
\end{problem}
\newpage 
\noindent {The problems so far are worth 100\% of the grade. Solving this problem provides extra 15\%. }
\begin{problem} (15 pts)
Compute the limit.
\[
\lim_{x\to \infty} \sqrt{ 3^{2x}+3^{x+1}}-\sqrt{3^{2x}-3^{x-1}}\quad .
\]

\end{problem}
\end{document}