\documentclass{article}
\usepackage{amsmath, amsfonts, amssymb, verbatim, hyperref}
\usepackage{auto-pst-pdf}
\usepackage{pst-plot}
\usepackage{multicol}
\addtolength{\hoffset}{-3.5cm}
\addtolength{\textwidth}{6.8cm}
\addtolength{\voffset}{-3.3cm}
\addtolength{\textheight}{6.3cm}
\renewcommand{\Re}{\mathrm{Re~}}
\renewcommand{\Im}{\mathrm{Im~}}
\newcommand{\doublebrace}[4]{\left\{\begin{array}{ll} #1 & #2 \\#3 & #4  \end{array} \right.}
\newcommand{\triplebrace}[6]{\left\{\begin{array}{ll} #1 & #2 \\#3 & #4  \\#5 & #6\end{array} \right.}
\date{}
\newtheorem{problem}{Problem}
\title{
Homework Math 140 \\
Lectures 5 and 6 \\
Will be quizzed Thursday Feb 21
}
\begin{document}
\maketitle
\textbf{This homework contains copyrighted material from  James Stewart, Calculus, 7th edition, 2012. 
You are not permitted to copy this file for any purpose other than completing your homework. You are not allowed to give a copy of this file to anyone outside of our course. 
}

\begin{problem}(Textbook page 70, problems 11-32). Complete problem 6 from the previous homework, if you failed to do so last week.
\end{problem}
\begin{problem}
(Textbook, page 91, problem 23-24). Find the (implied) domain of $f(x)$. Extend the definition of $f$ at $x=2$ to make $f$ continuous at $2$.
\begin{multicols}{2}
\begin{enumerate}
\item $f(x)=\frac{x^2-x-2}{x-2}$.
\item $f(x)=\frac{x^3-8}{x^2-4}$.
\end{enumerate}
\end{multicols}
\end{problem}
\begin{problem}(Textbook, page 92, problem 41-43)
Find the numbers $x$ for which $f$ is discontinuous. At which of these numbers is $f$ continuous from the right, from the left, or neither? 
\begin{multicols}{2}
\begin{enumerate}
\item $f(x)=\triplebrace{1+x^2}{\mathrm{if~} x\leq 0 } {2-x}{\mathrm{if~} 0<x\leq 2}{(x-2)^2}{\mathrm{if~} x>2}$.
\item $f(x)=\triplebrace{x+1}{\mathrm{if~} x\leq 1 }{\frac{1}{x}}{\mathrm{if~}1<x<3}{\sqrt{x-3}}{\mathrm{if~} x\geq 3}$.
\item $f(x)=\triplebrace{x+2}{\mathrm{if~} x<0}{2x^2}{\mathrm{if~} 0\leq x\leq 1 }{2-x}{\mathrm{if~}x>1 }$
\end{enumerate}
\end{multicols}
\end{problem}
\begin{problem} (Textbook, page 92, problem 46)
Find the values of $a$ and $b$ that make $f$ continuous everywhere.
\[f(x)=\triplebrace{\frac{x^2-4}{x-2}}{\mathrm{if~}x<2}{ax^2-bx+3}{\mathrm{if~}2\leq x<3}{2x-a+b}{x\geq 3} \quad .\]
\end{problem}
\begin{problem}
(Textbook, page 92, problem 51-56) Use the Intermediate Value Theorem to show that there is a real number solution of the given equation in the specified interval. 
\begin{multicols}{2}
\begin{itemize}
\item $x^4+x-3=0$ where $x\in (1,2)$.
\item $\sqrt[3]{x}=1-x$ where $x\in (0,2) $.
\item $\cos x=x$, where $x\in (0,1)$.
\item $\sin x=x^2-x$, where $x\in (1,2)$.
\item $\cos x=x^3$, where $x\in \mathbb R$ (i.e., $x$ is an arbitrary real number).
\item $x^5-x^2+2x+3$, where $x\in \mathbb R$.
\end{itemize}
\end{multicols}
\end{problem}
\begin{problem}(Textbook, page 93, problem 63-64) 
For which values of $x$ is $f$ continuous?
\begin{itemize}
\item $f(x)=\doublebrace{0}{\mathrm{if~} x\mathrm{~is~rational}}{1}{\mathrm{if~}x~\mathrm{is~irrational}}$
\item $f(x)=\doublebrace{0}{\mathrm{if~} x\mathrm{~is~rational}}{x}{\mathrm{if~}x~\mathrm{is~irrational}}$
\end{itemize}
\end{problem}
\begin{problem} This problem will not appear on the quiz, and might be slightly harder. For which values of $x$ is $f$ continuous?
\[f(x)=\doublebrace{\frac{1}{q^2}}{\mathrm{if~}x\mathrm{~is~rational~and~} x=\frac{p}{q} }{0}{\mathrm{if~}x~\mathrm{is~irrational}}\]
where in the first item $p,q$ are relatively prime integers (i.e., integers without a common divisor).
\end{problem}
\end{document}