\documentclass{article}
\usepackage{amsmath, amsfonts, amssymb, verbatim, hyperref}
\usepackage{auto-pst-pdf}
\usepackage{pst-plot}
\addtolength{\hoffset}{-3.5cm}
\addtolength{\textwidth}{6.8cm}
\addtolength{\voffset}{-3.3cm}
\addtolength{\textheight}{6.3cm}
\renewcommand{\Re}{\mathrm{Re~}}
\renewcommand{\Im}{\mathrm{Im~}}
\newcommand{\doublebrace}[4]{\left\{\begin{array}{ll} #1 & #2 \\#3 & #4  \end{array} \right.}

\date{}
\newtheorem{problem}{Problem}
\title{
Homework Math 140 \\
Lectures 1 and 2 \\
Will be quizzed Thursday Feb 7
}
\begin{document}
\maketitle
\begin{problem}
Visit \url{https://piazza.com/umb/spring2013/math140}. Download the .pdf file with the slides of Lecture 1. The address is 
\url{https://piazza.com/umb/spring2013/math140/resources}. Download the .pdf file with this homework from 
\url{https://piazza.com/umb/spring2013/math140/resources}. 

I strongly encourage you to register at the course. 
\url{https://piazza.com/umb/spring2013/math140}
Feel free to post questions or otherwise discuss the material at the above address.
\end{problem}

\begin{problem}
\begin{enumerate}
\item Evaluate and simplify $\frac{f(3+h)-f(3)}{h}$ as a function of $h$, where $f(x)= 1-x-3x^2$. 
\item Evaluate and simplify $\frac{f(a+h)-f(a)}{h}$ as a function of $h$, where $f(z)= 1-z-3z^2$ and $a$ is a parameter.
\item Evaluate and simplify $\frac{f(x)-f(a)}{x-a}$ as a function of $x$, where $f(x)= \frac{1}{x}$ and $a$ is a parameter.
\item Evaluate and simplify $\frac{f(x)-f(2)}{x-2}$ as a function of $x$, where $f(x)= \frac{x+1}{x+2}$.
\end{enumerate}
\end{problem}
\begin{problem}Find the implied domain of the function 
\begin{enumerate}
\item $f(x)=\frac{x+4}{x^2+3x-4}$.
\item $g(t)=\sqrt{5-t}-\sqrt{3+t}$.
\item $f(x)=\sqrt{3-\sqrt{3-x}}$.
\end{enumerate}
\end{problem}
\begin{problem}
Plot the piecewise defined functions.
\begin{enumerate}
\item $f(x)=\doublebrace{x^2}{x\in (-\infty,2)}{x+2}{x\geq 2}$
\item $f(x):=|3-5x|$.
\end{enumerate}
\end{problem}
\begin{problem}
Write down formulas for function whose graphs are as follows. The graphs are up to scale. The arc is a part of a circle.
\begin{enumerate}
\item 
\begin{pspicture}(0,0)(3,3)
\psaxes{->}(0,0)(0,0)(3,3)
\psline[linecolor=red](0,0)(1,2)
\psline[linecolor=red](1,2)(2,0.5)
\psline[linecolor=red](2,0.5)(3,0)
\rput[c](2,0.5){$\bullet$}
\rput[bl](2,0.5){$(2,1/2)$}
\end{pspicture}
\item \begin{pspicture}(0,0)(3,3)
\psaxes{->}(0,0)(0,0)(3,3)
\psplot[linecolor=red]{0}{2}{x 2 mul x x mul sub sqrt }
\psline[linecolor=red](2,0)(3,2)
\end{pspicture}
\end{enumerate}
\end{problem}
\begin{problem} Find the functions $f\circ g$, $g\circ f$, $f\circ f$ and $g\circ g$ and their implied domains.
\begin{enumerate}
\item $f(x)=x^2+1$, $g(x)=x+1$.
\item $f(x)=\sqrt{x+1}$, $g(x)=x+1$.
\item $f(x)= 2x$, $g(x)= \tan x$.
\item $f(x)=\frac{x+1}{x-1}$, $g(x)=\frac{x-1}{x+1}$.
\end{enumerate}
\end{problem}
\begin{problem}Graph the functions by hand, by applying consecutively the transformations learned in class.
\begin{itemize}
\item $y=\frac{1}{x}$.
\item $y=\frac{1}{x+1}$.
\item $y=\frac{1}{2x+1}$.
\item $y=\frac{3}{2x+1}$.
\item $y=\frac{3+x}{2x+1}$.
\item $y=\left|\frac{3+x}{2x+1}\right|$.
\end{itemize}
\end{problem}
\end{document}