\documentclass{article}
\usepackage{../homework-problems-UMB}

\toggletrue{solutions}
%\togglefalse{solutions}
\toggletrue{answers}

\begin{document}
\begin{center}
\Large
Review problems Test 1\\ Math 140 Calculus I \\ \normalsize Instructor: Todor Milev
\end{center}

%\noindent \textbf{Name:} \hfill{~}
%\begin{tabular}{c|c|c|c|c|c|c|c|c||c}
%Problem&1 &2&3&4&5&6&7&8& $\sum$\\ \hline
%Score&&&&&&&&&\\ \hline
%Max&20&20&20&20&20&10&20&20&150
%\end{tabular} 

\noindent The exam is closed books, no calculators will be allowed. The time for work will be 50 minutes. The problems on the exam will be similar to the problems in the review sheet. You will be asked a theoretical question (see the last two problems).

\begin{enumerate}
\item \input{../../modules/precalculus/homework/functions-composing-fractional-linear2}
\item \input{../../modules/trigonometry/homework/trigonometry-equations2.tex}
\item \input{../../modules/limits/homework/limit-x-tends-to-a-common-factor2}
\input{../../modules/curve-sketching/homework/asymptotes-vertical2}

\item \input{../../modules/continuity/homework/IVT-problems2}
\item \input{../../modules/curve-sketching/homework/asymptotes-vertical-horizontal2}
\item \input{../../modules/curve-sketching/homework/asymptotes-vertical-horizontal3}
\item 
\begin{enumerate}
\item State the Intermediate Value Theorem. (See Lecture 5).
\item Give the $\varepsilon,\delta$-definition of limit. (You are asked to reproduce the sentence that involves the symbols $\varepsilon,\delta$ from Lecture 4).
\end{enumerate}
\end{enumerate}
\end{document}





\end{document}