% begin module natural-logarithm-def
\begin{frame}
\alertNoH{1,2}{What does $\log x$ stand for?} \uncover<2->{ \alertNoH{2}{\textbf{WARNING:}} there are  \alertNoH{2}{ \textbf{two different}} accepted uses for $\log x$.}
\vskip -0.15cm
\begin{columns}
\column{0.4\textwidth}
\begin{itemize}
\item<3-> In some texts/applications $\log x$ stands for
\[
\log x = \log_{10} x\quad .
\]
\item<4-> Used in many engineering texts.
\item<4-> Used in many natural sciences texts.
\item<4-> Used in many high school textbooks.
\item<4-> Used in old math textbooks.
\end{itemize}
\column{0.63\textwidth}
\begin{itemize}
\item<5-> In other texts/applications $\log x$ stands for {\color{gray}(the principal branch of the)} \textbf{complex logarithm}

$
\alertNoH{7}{\log x =}\left\{\begin{array}{ll} \alertNoH{7}{\ln x} =\log_e x& \alertNoH{7}{ \text{if  } x>0} \\ \ln (-x) +\pi i&\text{if } x<0 \\ \textbf{?} &\text{for } x \notin \mathbb R\end{array}\right. \quad .
$
\begin{itemize}
\item<6-> Used in mathematical, many computer science texts.
\item<6-> Used in many natural science texts.
\item<6-> Used in most computer algebra systems.
\item<6-> This is the notation accepted by most mathematicians.
\end{itemize}
\item<7-> $\log $ and $\ln $ have different domains but else coincide: $\ln$ is defined for positive reals, and $\log$ - for non-zero complex.
\end{itemize}
\end{columns}
\end{frame}

\begin{frame}
\begin{emptyTheorem}
\begin{itemize}
\item \alert<1->{In the present course we shall abstain from using the notation $\log x$. }
\item When we need logarithms base $10$ we will always write $\log_{10}$.
\end{itemize}
\end{emptyTheorem}

\begin{itemize}
\item Within this course, we request that the student abstain from using $\log x$ and use instead the unambiguous $\log_{10} x$.
\item Outside of this course, we recommend that the student continue avoiding the notation $\log$.
\item Should our recommendation contradict the commonly accepted conventions in the field of study of the student, we expect the student to honor the conventions of their fields of study.
\end{itemize}
\end{frame}
% end module natural-logarithm-def
