\documentclass[12pt]{article}
\usepackage{amsmath}
%\usepackage{cancel}

%compiles with pdflatex -shell-escape
\input{../homework-problems-UMB}

\toggletrue{solutions}
\toggletrue{answers}

\newtheorem{problem}{Problem}
\renewcommand{\fcProblemRef}{\theproblem.\theenumi}
\renewcommand{\fcSubProblemRef}{\theenumi.\theenumii}

\newcommand{\finishUMBExamPage}{
\vfill
%\begin{tabular}{l}
%****CONFIDENTIAL **** \\ **** PLEASE DO NOT PRINT  **** 
%\end{tabular}
\hfill \begin{tabular}{|c|}\hline Score\\\hline  \\\hline \end{tabular}
}
\newcommand{\startUMBExamPage}{\newpage
\noindent Name: \underline{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\hfill ID:\underline{~~~~~~~~~~~~~~~}}

\pagestyle{empty}
\begin{document}
\begin{center}
\Huge 
\Huge
\textsc{Math 140 Final Exam}\\
\textsc{Solutions}\\
\Large Spring 2016\\
University of Massachusetts Boston
\end{center}
\textbf{Instructions.}
\begin{enumerate}
\item Write your name on every page.
\item This is a closed book exam.
\item Show your answers and arguments for your answers in the space provided.
\item Answers without proper justification will \textbf{not receive full credit}.
\item \textbf{Calculators, cell phones and other electronic devices are not allowed.}
\item Grading table. 

{\Large
\begin{tabular}{|c|c|c|}\hline
Problem & Points & Score\\\hline\hline
1& 15 &\\\hline
2& 10 &\\\hline
3& 25 &\\\hline
4& 15 &\\\hline
5& 10 &\\\hline
6& 10 &\\\hline
7& 10 &\\\hline
8&  15 &\\\hline
9&  10 &\\\hline
10& 10 &\\\hline
11& 10 &\\\hline
$\sum$ & 140 &\\\hline
\end{tabular}
}
\end{enumerate}
\begin{problem}
Evaluate the limit or show that it does not exist.
\begin{enumerate}[ref={\fcProblemRef}]
\item 
\input{../../modules/limits/homework/limit-x-tends-to-a-common-factor-rational-function-subproblem-2-1}
\item \input{../../modules/limits/homework/limit-x-tends-to-a-difference-quotient-subproblem-23}
\item \input{../../modules/limits/homework/limit-x-tends-to-infinity-1-problem-10-1}
\end{enumerate}
\input{../../modules/limits/homework/limit-x-tends-to-a-common-factor-rational-function-subproblem-2-1-solution}
\input{../../modules/limits/homework/limit-x-tends-to-a-difference-quotient-subproblem-23-solution}
\input{../../modules/limits/homework/limit-x-tends-to-infinity-1-problem-10-1-solution}


\end{problem}
\begin{problem}
Find the vertical and horizontal asymptotes of the function.

\input{../../modules/curve-sketching/homework/asymptotes-vertical-horizontal-1-subproblem-10}
\end{problem}
\input{../../modules/curve-sketching/homework/asymptotes-vertical-horizontal-1-subproblem-10-solution}

\begin{problem}
Differentiate with respect to $x$.
\begin{enumerate}[ref={\fcProblemRef}]
\item \input{../../modules/derivatives/homework/derivatives-trig-2-problem-8-1}
\item \input{../../modules/derivatives/homework/derivatives-product-quotient-rule-3-problem-4-1}
\item \input{../../modules/chain-rule/homework/chain-rule1-problem-16}
\item \input{../../modules/chain-rule/homework/chain-rule1-problem-2-1}
\item \input{../../modules/logarithms/homework/derivatives-involving-logarithms-1-problem-15}
\end{enumerate}
\end{problem}
\input{../../modules/derivatives/homework/derivatives-trig-2-problem-8-1-solution}
\input{../../modules/derivatives/homework/derivatives-product-quotient-rule-3-problem-4-1-solution}
\input{../../modules/chain-rule/homework/chain-rule1-problem-16-solution}
\input{../../modules/chain-rule/homework/chain-rule1-problem-2-1-solution}
\input{../../modules/logarithms/homework/derivatives-involving-logarithms-1-problem-15-solution}


\begin{problem}
Verify that the coordinates of the given point satisfy the given equation. Use implicit differentiation to find an equation of the tangent line to the curve at the given point. 

\input{../../modules/implicit-differentiation/homework/implicit-tangent-problem-5}
\end{problem}
\input{../../modules/implicit-differentiation/homework/implicit-tangent-problem-5-solution}

\begin{problem}Find the maximum and the minimum of the function over the indicated closed interval. Indicate for which values of $x$ is the maximum/minimum attained.
\begin{enumerate}[ref={\fcProblemRef}]
\item  \input{../../modules/maxima-minima/homework/maxima-minima-problem-4-1}
\item \input{../../modules/maxima-minima/homework/maxima-minima-problem-16}
\end{enumerate}
\end{problem}
\input{../../modules/maxima-minima/homework/maxima-minima-problem-4-1-solution}
\input{../../modules/maxima-minima/homework/maxima-minima-problem-16-solution}
\begin{problem}
\input{../../modules/curve-sketching/homework/sketch-graph-with-all-the-details-2-problem-3}
\end{problem}
\input{../../modules/curve-sketching/homework/sketch-graph-with-all-the-details-2-problem-3-solution}

\begin{problem}
{\sc Do not solve the integral. } Compute a \textbf{Riemann sum} for the indicated integral using the indicated number of intervals with the indicated sampling point. 
\input{../../modules/integration/homework/riemann-sum-problems-1-problem-8}
\end{problem}
\input{../../modules/integration/homework/riemann-sum-problems-1-problem-8-solution}


\begin{problem}
Integrate.
\begin{enumerate}[ref={\fcProblemRef}]
\item \input{../../modules/integration/homework/basic-definite-integrals-2-problem-3}
\item  \input{../../modules/substitution-rule/homework/substitution-rule-definite-1-problem-5}
\item \input{../../modules/substitution-rule/homework/substitution-rule-2-problem-13}
\end{enumerate}
\end{problem}
\input{../../modules/integration/homework/basic-definite-integrals-2-problem-3-solution}
\input{../../modules/substitution-rule/homework/substitution-rule-definite-1-problem-5-solution}
\input{../../modules/substitution-rule/homework/substitution-rule-2-problem-13-solution}
\begin{problem}
Find the  derivative of $f(x)$ using the Fundamental Theorem of Calculus, Part I.
\begin{enumerate}[ref={\fcProblemRef}]
\item \input{../../modules/integration/homework/FTC-part1-problems-1-problem-5}
\item 
\input{../../modules/integration/homework/FTC-part1-problems-1-problem-6}
\end{enumerate}
\end{problem}
\input{../../modules/integration/homework/FTC-part1-problems-1-problem-5-solution}
\input{../../modules/integration/homework/FTC-part1-problems-1-problem-6-solution}
\begin{problem}~
\input{../../modules/area-between-curves/homework/area-between-curves-problems-1-problem-4}
\end{problem}
\input{../../modules/area-between-curves/homework/area-between-curves-problems-1-problem-4-solution}

\begin{problem}
\input{../../modules/volumes/homework/solids-of-revolution-problems-1-problem-4}
\end{problem}
\input{../../modules/volumes/homework/solids-of-revolution-problems-1-problem-4-solution}


\end{document}
