\documentclass[12pt]{article}
\usepackage{amsmath}
%\usepackage{cancel}

%compiles with pdflatex -shell-escape
\input{../../lectures/pstricks-commands}


\addtolength{\hoffset}{-2.5cm}
\addtolength{\textwidth}{4.8cm}
\addtolength{\voffset}{-3cm}
\addtolength{\textheight}{6cm}

\newtheorem{problem}{Problem}
\newcommand{\diff}{{\normalfont \text{d}}}

\newcommand{\finishUMBExamPage}{
\vfill
%\begin{tabular}{l}
%****CONFIDENTIAL **** \\ **** PLEASE DO NOT PRINT  **** 
%\end{tabular}
\hfill \begin{tabular}{|c|}\hline Score\\\hline  \\\hline \end{tabular}
}
\newcommand{\startUMBExamPage}{\newpage
\noindent Name: \underline{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\hfill ID:\underline{~~~~~~~~~~~~~~~~~~~~~~~}}

\pagestyle{empty}
\begin{document}
\begin{center}
\Huge 
\Huge
%****CONFIDENTIAL **** \\ **** PLEASE DO NOT PRINT  **** \\
\textsc{Math 140 Final Exam}\\
\Large Spring 2016\\
University of Massachusetts Boston
\end{center}

\medskip

\noindent Name: \underline{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\hfill ID:\underline{~~~~~~~~~~~~~~~~~~~~~~~}

\bigskip

\bigskip


\textbf{Instructions.}
\begin{enumerate}
\item Write your name on every page.
\item This is a closed book exam.
\item Show your answers and arguments for your answers in the space provided.
\item Answers without proper justification will \textbf{not receive full credit}.
\item \textbf{Calculators, cell phones and other electronic devices are not allowed.}
\end{enumerate}

\hrulefill \textsc{Do not write below this line. }\hrulefill

\vfill
{\Large
\begin{tabular}{|c|c|c|}\hline
Problem & Points & Score\\\hline\hline
1& 15 &\\\hline
2& 10 &\\\hline
3& 25 &\\\hline
4& 15 &\\\hline
5& 10 &\\\hline
6& 10 &\\\hline
7& 10 &\\\hline
8&  15 &\\\hline
9&  10 &\\\hline
10& 10 &\\\hline
11& 10 &\\\hline
$\sum$ & 140 &\\\hline
\end{tabular}
}
\startUMBExamPage
\begin{problem}
Evaluate the limit or show that it does not exist.
\begin{itemize}
\item $\displaystyle \lim_{x\to -2} \frac{2x^2+x-6}{x^2-4}$

\vfill
\item $ \displaystyle \lim_{h\to 0} \frac{\frac{1}{(2+h)^2}-\frac{1}{4}}{h}  $.

\vfill
\item $ \displaystyle \lim_{x\to \infty} \frac{\sqrt{3x^2+2x+1}}{x+1} $. 
\end{itemize}


\end{problem}
\finishUMBExamPage
\startUMBExamPage
\begin{problem}
Let 
\[
f(x)= \frac{x}{\sqrt{x^2+3} -2x}
\]
\begin{itemize}
\item Find the vertical asymptotes of $f(x)$.
\item Find the horizontal asymptotes of $f(x)$. Make sure you investigate both directions on the $x$ axis.
\end{itemize}
\end{problem}
\finishUMBExamPage
\startUMBExamPage
\begin{problem}
Differentiate the function with respect to $x$.
\begin{itemize}
\item $\displaystyle f(x)=(\sec x )e^{x}$.

\vfill
\item $\displaystyle f(x)= \frac{x+1}{x^3+1}$.
\vfill
\newpage
\item    $\displaystyle f(x)={\scalebox{1.4}{ e}} 
^{{\raisebox{1ex}{$-\frac{1}{x}$}}}
$.

\vfill\item $f(x)=\sqrt{1-\sqrt{x}}$.

\vfill
\item $f(x)=\ln (\sec x) +\ln (\cot x)$.

\end{itemize}
\end{problem}
\finishUMBExamPage

\startUMBExamPage


\begin{problem}
Let $x$ and $y $ be related via $\displaystyle y^{3}+x^{3}+4 x y=\frac{3}{4}$. The curve given by the points that satisfy this equation is plotted below.

\psset{xunit=0.7cm, yunit=0.7cm}
\begin{pspicture}(-2,-2)(2,2)
\fcAxesStandard{-2}{-2}{2}{2}
\fcImplicitIId[linecolor=red, linestyle=dashed, dashes={[1 1] 0}, showGridImplicitIId=false, useMidpointImplicitPlots=false]{-3}{-3}{120}{120}{0.05}{0.05}{y y y mul mul x x x mul mul add 4 x y mul mul add 0.75 sub} 
\fcFullDot{-0.5}{-0.5}
\end{pspicture}



\begin{itemize}
\item Using implicit differentiation, express the derivative $\frac{\diff y}{\diff x} $ in terms of  $x$ and $y$.
\item Show that the point $\left(-\frac{1}{2},-\frac{1}{2}\right)$ lies on the curve. 
\item Find the equation of the tangent line to that curve at the point $\left(-\frac{1}{2},-\frac{1}{2}\right)$.
\end{itemize}

\end{problem}
\finishUMBExamPage
\startUMBExamPage
\begin{problem}Find the maximum and the minimum of the function over the indicated closed interval. Indicate for which values of $x$ is the maximum/minimum attained.
\end{problem}
\begin{itemize}
\item $f(x)=x^3-x+1$ over the interval $[-2,1]$.
\vskip 10cm 

\item $f(x)=x e^{2x}$ over the interval $\left[ -2,\frac{1}{2}\right]$.


\end{itemize}
\finishUMBExamPage
\startUMBExamPage
\begin{problem}
Consider the function $\displaystyle f(x)=\frac{4 x^{2}+10 x+5}{2 x+1} $. Computation shows that $\displaystyle f'(x)=\frac{8 x^{2}+8 x}{\left(2 x+1\right)^{2}}$ and $\displaystyle f''(x)=\frac{8}{\left(2 x+1\right)^{3}} $.

\begin{itemize}
\item Find the intervals of increase and intervals of decrease of $f$.
\vfill 
\item Find the local maxima and minima of $f$. 
\vfill 
\item Find where the function is concave up and where it is concave down.
\vfill 
\newpage 
\item Sketch the function $f(x)$ roughly by hand. Make sure that your plot matches your computations from the preceding parts of the problem.


You may use the provided grid and coordinate system. From the previous page, we recall that $\displaystyle f(x)=\frac{4 x^{2}+10 x+5}{2 x+1} $, $\displaystyle f'(x)=\frac{8 x^{2}+8 x}{\left(2 x+1\right)^{2}}$ and $\displaystyle f''(x)=\frac{8}{\left(2 x+1\right)^{3}} $.

The 4 points plotted on the grid are known to lie on the curve.

\psset{xunit=1cm, yunit=1cm}
\begin{pspicture}(-3,-3)(3,12.1)
\fcAxesStandard{-3}{-3}{3}{12}
\fcGrid[linestyle=dashed, linewidth=0.5, linecolor=gray]{-3}{-3}{6}{15}{1}{1}{}
\rput[t](0.9,-0.2){$1$}
\fcLabels{3}{12}
\fcFullDot{-3}{1 dict begin /x -3 def x x mul 4 mul 10 x mul 5 add add 2 x mul 1 add div end}
\fcFullDot{-0.59}{1 dict begin /x -0.59 def x x mul 4 mul 10 x mul 5 add add 2 x mul 1 add div end}
\fcFullDot{-0.44}{1 dict begin /x -0.44 def x x mul 4 mul 10 x mul 5 add add 2 x mul 1 add div end}
\fcFullDot{3}{1 dict begin /x 3 def x x mul 4 mul 10 x mul 5 add add 2 x mul 1 add div end}

%\psplot{-3}{-0.59}{x x mul 4 mul 10 x mul 5 add add 2 x mul 1 add div}
%\psplot{-0.44}{3}{x x mul 4 mul 10 x mul 5 add add 2 x mul 1 add div}
\end{pspicture}

\end{itemize}
\end{problem}


\finishUMBExamPage
\startUMBExamPage
\begin{problem}
%\begin{itemize}
%\item 
{\sc Do not solve the integral. } Compute a \textbf{Riemann sum} for the indicated integral using \textbf{$3$ intervals}. Choose the sampling points to be the \textbf{left endpoints} of each interval. 
\[
\int_{-1}^0\frac{1}{3 {{x}}^{2}+1}\diff x
\]
Simplify your answer to a rational number (single fraction of two integers). {\sc Non-simplified answers will not receive full credit.}

\vfill
%\item Express the differential $\diff \left(\sqrt{x}\right)$ using the differential $\diff x$.
\vfill
%\end{itemize}
\end{problem}
\finishUMBExamPage
%\startUMBExamPage
%\begin{problem}
%Let $f$ be a continuous function for which $\int_{1}^3 f(x)\diff x=8$. Show that there exists a point $c$ between $1$ and $3$ such that $f(c)=4$.	
%\end{problem}
%\finishUMBExamPage
\startUMBExamPage
\begin{problem}
Integrate.
\begin{itemize}
\item $\displaystyle\int \frac{\sqrt[3]{x}-x^{\frac{1}{2}}+1 }{x}\diff x$.
\vskip 5cm
\item $\displaystyle\int_{-3}^{-2} \frac{x}{1-x^2}\diff x$.
\vskip 5cm
\item $\displaystyle\int \frac{\sin (\ln x)}{x} \diff x$.
\end{itemize}
\end{problem}
\finishUMBExamPage
\startUMBExamPage
\begin{problem}
Find the indicated derivatives using the Fundamental Theorem of Calculus, Part I.
\begin{itemize}
\item $\displaystyle \frac{\diff}{\diff x}\int_{1}^x\left(\sqrt{t}-\sqrt[3]{t}\right)\diff t$.

\vskip 8cm
\item $\displaystyle \frac{\diff}{\diff x}\int_{1}^{\frac{1}{x+1} }\sin \left(  t^2\right) \diff t$.
\end{itemize}
\end{problem}

\finishUMBExamPage
\startUMBExamPage
\begin{problem}
\begin{itemize}
\item Sketch the region bounded by the curves $y=x^2$ and $y=2x^2+x-2$.

\psset{xunit=0.5cm, yunit=0.5cm}
\begin{pspicture}(-3,-3)(3,3)
\fcAxesStandard{-3.5}{-3.5}{2.5}{5.5}
\fcGrid[linestyle=dashed, linewidth=0.5, linecolor=gray]{-3}{-3}{5}{8}{1}{1}{}
\rput[t](0.9,-0.2){$1$}
\fcLabels{3.5}{5.5}
%\psplot{-3}{2}{x x mul}
%\psplot{-3}{2}{x x mul 2 mul x -2 add add}
\end{pspicture}

\vskip 2cm


\item Find the area of the region.
\end{itemize}
\end{problem}
\finishUMBExamPage
\startUMBExamPage
\begin{problem}
Set up \textsc{but do not evaluate} an integral to calculate the volume of the solid obtained by rotating the region bounded by $y=-x^2+2$ and $y=0$ about the given line. \textsc{Credit will be awarded only for a correct integral, you do not need to solve it.}

\begin{itemize}
\item The $x$ axis.

\vskip 8cm
\item The line $y=-3$.
\end{itemize}
\end{problem}

\finishUMBExamPage
\end{document}
