\documentclass{article}
\input{../homework-problems-UMB}

\toggletrue{solutions}
\toggletrue{answers}
\renewcommand{\fcProblemRef}{\theproblem.\theenumi}
\renewcommand{\fcSubProblemRef}{\theproblem.\theenumi.\theenumii}


\newcommand{\hide}[1]{}
\newtheorem{problem}{Problem}
\pagestyle{empty}
\begin{document}
\begin{center}
\Large
Review sheet Test 2 \\ Math 140 Calculus I \\ \normalsize Spring 2016 \\ Instructor(s): Todor Milev, Ping Xu
\end{center}
%\noindent \textbf{Name:\underline{~~~~~~~~~~~~~~~~~~~~~~~} } \hfill{~}



\noindent The exam is closed textbook. \textbf{No electronic devices are allowed during the exam. } The exam will contain a number of problems and sub-problems (the number of problems may depend on your section/instructor). \textbf{Expect each problem type} from this review sheet to appear on the test. Some exam problems \textbf{may be formulated in a slightly different fashion.}%You are allowed one single formula sheet, handwritten by you. No template problem solutions are allowed. The sheet will be collected with the test. Photocopied formula sheets are not allowed. 

\begin{problem}Differentiate.
\begin{multicols}{2}
\begin{enumerate}[ref={\fcProblemRef}]
\item \input{../../modules/derivatives/homework/derivatives-product-quotient-rule-2-problem-6}
\item \input{../../modules/derivatives/homework/derivatives-product-quotient-rule-2-problem-7}
\item \input{../../modules/derivatives/homework/derivatives-product-quotient-rule-2-problem-11}
\item \input{../../modules/derivatives/homework/derivatives-trig-problem-2}
\item \input{../../modules/derivatives/homework/derivatives-trig-problem-6}
\item \input{../../modules/derivatives/homework/derivatives-trig-problem-14}
\item 
\input{../../modules/chain-rule/homework/chain-rule1-problem-13}
\item 
\input{../../modules/chain-rule/homework/chain-rule1-problem-14}
\item 
\input{../../modules/chain-rule/homework/chain-rule1-problem-15}
\item 
\input{../../modules/logarithms/homework/derivatives-involving-logarithms-1-problem-7}
\item 
\input{../../modules/logarithms/homework/derivatives-involving-logarithms-1-problem-10}
\item \input{../../modules/logarithms/homework/derivatives-involving-logarithms-1-problem-11}

\end{enumerate}
\end{multicols}
\end{problem} 
\input{../../modules/derivatives/homework/derivatives-product-quotient-rule-2-problem-11-solution}
\input{../../modules/logarithms/homework/derivatives-involving-logarithms-1-problem-11-solution}
\begin{problem}Verify that the coordinates of the given point satisfy the given equation. Use implicit differentiation to find an equation of the tangent line to the curve at the given point.

\begin{enumerate}[ref={\fcProblemRef}]
\item \input{../../modules/implicit-differentiation/homework/implicit-tangent-problem-1}
\item \input{../../modules/implicit-differentiation/homework/implicit-tangent-problem-3}
\item
\input{../../modules/implicit-differentiation/homework/implicit-tangent-problem-4}

\end{enumerate}
\end{problem}
\input{../../modules/implicit-differentiation/homework/implicit-tangent-problem-1-solution}

\begin{problem}
Find the maximum and minimum values of $f$ on the given interval. 
\begin{multicols}{2}
\begin{enumerate}[ref={\fcProblemRef}]
\item \input{../../modules/maxima-minima/homework/maxima-minima-problem-2}
\item \input{../../modules/maxima-minima/homework/maxima-minima-problem-3}
\item \input{../../modules/maxima-minima/homework/maxima-minima-problem-4}
\end{enumerate}
\end{multicols}
\end{problem}
\begin{problem}
\input{../../modules/maxima-minima/homework/mvt-show-only-one-root-1}
\end{problem}
\input{../../modules/maxima-minima/homework/mvt-show-only-one-root-1-solutions}
\begin{problem}
Find the
\begin{multicols}{2}
\begin{itemize}
\item the implied domain of $f$,
\item $x$ and $y$ intercepts of $f$,
\item horizontal and vertical asymptotes,
\item intervals of increase and decrease,
\item local and global minima, maxima,
\item intervals of concavity,
\item points of inflection.
\end{itemize}
\end{multicols}
Label all relevant points on the graph. Show all of your computations.
\begin{enumerate}[ref={\fcProblemRef}]
\item \input{../../modules/curve-sketching/homework/sketch-graph-with-all-the-details-problem-2}
\item \input{../../modules/curve-sketching/homework/sketch-graph-with-all-the-details-problem-7}
\end{enumerate}
\end{problem}
\input{../../modules/curve-sketching/homework/sketch-graph-with-all-the-details-problem-2-solution}



\end{document}