\documentclass{article}
\usepackage{amsmath, amsfonts, amssymb, verbatim, hyperref}
\usepackage{enumitem}
\usepackage{pst-plot}
\usepackage{pstricks}
\usepackage{lscape}

\addtolength{\hoffset}{-3.5cm}
\addtolength{\textwidth}{6.8cm}
\addtolength{\voffset}{-3.3cm}
\addtolength{\textheight}{6.3cm}

\newcounter{topicsCounter}
\newcounter{topicsSubCounter}[topicsCounter]
\newcounter{topicsSubSubCounter}[topicsSubCounter]

\newcommand{\skipped}{\begin{tabular}{@{}c}optional \\reading \\not on test \end{tabular}}

\usepackage{longtable}
\usepackage{xr}
%\externaldocument{../../homework/UMB-All-Problems-By-Course/Calc-I-MasterProblemSheet}
\externaldocument{./Calc-I-MasterProblemSheetOneFile}

\newcommand{\refBad}[1]{%
\ifthenelse{\equal{\ref{#1}}{??}}%
{(n/a)}%
{\ref{#1}}%
}%example usage: \refBad{\ref{eqMacLaurinDef}}{their definition}{their definition (Definition \ref{eqMacLaurinDef})}


\newcommand{\counterTopic}{\refstepcounter{topicsCounter}\thetopicsCounter. }
\newcommand{\counterSubTopic}{\refstepcounter{topicsSubCounter}\thetopicsCounter.\thetopicsSubCounter.  }
\newcommand{\counterSubSubTopic}{\refstepcounter{topicsSubSubCounter}\thetopicsCounter.\thetopicsSubCounter.\thetopicsSubSubCounter. }

\newcommand{\apex}{A\kern -1pt \lower -2pt\hbox{P}\kern -4pt \lower .7ex\hbox{E}\kern -1pt X}


\newcommand{\websitebase}{https://piazza.com/umb/fall2015/math140}

\usepackage{pdfpages}

\title{Math 140 Calculus I \\ Fall 2015}
\date{}
\begin{document}

%\color{green}
\maketitle
%\noindent\textbf{Time and place.}
%Monday, Wednesday, Friday 10-10:50, McCormack, Room 417, first floor. Monday 11:00-11:50,

\noindent \textbf{Instructors.} 
\begin{tabular}{ll}
Todor Milev & \href{mailto:todor.milev@gmail.com}{\nolinkurl{todor.milev@gmail.com}} \\
Ping Xu & \href{mailto:pingxu0126@gmail.com}{\nolinkurl{pingxu0126@gmail.com}} 
\end{tabular}

\medskip
\noindent \textbf{Office hours. } \begin{tabular}{lp{12cm}}
Todor Milev & By appointment, or walk-in office hours: Tuesday 12:30-14:00, Thursday 15:30-17:00 (if I am not in the office I may be off to lunch).  Room: S-03-65.\\
Ping Xu&  To be announced.
\end{tabular}





\medskip \noindent \textbf{Lecture slides. }  \url{\websitebase/resources}

\medskip\noindent The lecture slides may be updated as the course progresses.


\medskip \noindent \textbf{Master Problem Sheet. }  \url{\websitebase/resources} 

\medskip\noindent The master problem sheet contains a collection of Calculus I problems. 

\medskip
\noindent \textbf{Homework.} You will be assigned homework, which will be posted on

\url{\websitebase/resources} \quad \quad \quad .

\noindent You will be expected to complete the homework in written form in a convenient for you format (notebook, folder, etc.). However, \textbf{the instructor(s) will not check/collect/proofread your homework.} 
 
\medskip
\noindent \textbf{Quizzes.} You will be given quizzes in class. \textbf{The time of the quiz will be announced in class}. Quizzes may be announced from one lecture day to the next. Your quiz problem will be one of your homework problems, verbatim (no number changes).





\medskip\noindent \textbf{Textbook. } There are two official textbooks for this course. It is  \textbf{mandatory} to have access to \textbf{at least one of the textbooks}. Having only one of the two textbooks is sufficient to complete the course. No mandatory homework will be assigned from either textbook. 

\begin{itemize}
\item Option I. The textbook \apex{} 3.0, Chapters 1-7. To download a free pdf file of the textbook, or to buy a physical copy of the textbook, visit the following site.

\url{http://www.apexcalculus.com/downloads/} 
\item Option II. James Stewart, Calculus, $7^{th}$ or $8^{th}$ edition (both editions are acceptable).
\end{itemize}

%\medskip
%\noindent \textbf{Prerequisite. } A standard pre-calculus course or equivalent.


\medskip
\noindent \textbf{Grades.} Your grade will consist of two tests, a comprehensive final exam, and a number of quizzes. 
\begin{itemize}
\item The quizzes will account for 20\% of your total grade.
\item The tests will account for 50\% (25\% each) of your total grade.
\item The final will account for 30\% of your total grade.
\end{itemize}
Please note that missed tests can not be made up, unless there is a valid medical reason accompanied with an official signed document from a medical doctor. Letter grades will be assigned as follows. 

\begin{center}
\begin{tabular}{lc|lc}
A & 85-100 & C & 65-69 \\
A-& 82-84 & C- & 62-64 \\
B+& 80-81 & D+ & 60-61 \\
B & 75-79& D & 55-59\\
B-& 72-74& D- & 50-54\\
C+& 70-71& F & below 50\\
\end{tabular}

\end{center}

No books, notes, calculators or any other electronic device (such as mobile phones) are allowed during any exam unless otherwise stated.

\medskip
\noindent \textbf{Student conduct.} Students  are required to adhere the University Policy on Academic Standards and Cheating, to the University Statement of Plagiarism and the Documentation of Written Work, and to the Code of Student Conduct as described in the catalog of Undergraduate programs, pages 44-45 and 48-52. The code is available at the following web-page.

\noindent\url{http://www.umb.edu/life_on_campus/policies/code/}


\begin{landscape}

\noindent \begin{longtable}{|@{}r@{}l@{}l@{~}l@{~}c cccc|}\hline
\multicolumn{9}{|c|}{\textbf{List of topics.} The list of topics is a preliminary guideline, and will be subject to change.
}\\\hline
&&& Topic & \apex{} 3.0 textbook & \begin{tabular}{l}Stewart, \\ Calculus, $7^{th}$ ed. \end{tabular} & \begin{tabular}{l} Relevant problems \\ Master Problem\\ Sheet\end{tabular} & \begin{tabular}{l}Lecture  \end{tabular} & \begin{tabular}{l} Expected \\ Week \\ (total 14) \end{tabular}  \\\hline
\counterTopic &&& Functions (Brief review).  & N/A (lectures only) & Chapter 1.1-1.3 & Section \refBad{secMPSfunctionBasics} &  \begin{tabular}{l}1,2\\ some \\ slides \\skipped\end{tabular} & 1 \\ \hline 
%Ways to represent a function.
%Some essential functions.
%New functions from old.
\counterTopic &&& Trigonometry. (Brief review). & N/A(lectures only) & N/A (lectures only)& Section \refBad{secMPStrigonometry}& \begin{tabular}{l}3\\ some \\ slides \\skipped\end{tabular} & 1  \\\hline
%Angles.
%The trigonometric functions.
%Trigonometric identities.
%Trigonometric identities, complex numbers and Euler's formula.
%Graphs of the trigonometric functions.
\counterTopic&&& Limits and continuity.&&&&& \\
&\counterSubTopic && Limits. One-sided limits. & Chapter 1.1, 1.2,1.4& Chapter 1.5, 1.7& N/A(theory only) & 4& 2 \\
&\counterSubTopic && The limit laws.& Chapter 1.3 & Chapter 1.6& Section \refBad{secMPSlimitsXtendsToNumer} & 4& 2 \\
&\counterSubTopic && Continuity. & Chapter 1.5  & Chapter 1.8 & Section \refBad{secMPScontinuityPiecewise} & 5& 2\\
&\counterSubTopic && Intermediate Value Theorem. & Chapter 1.5  & Chapter 1.8 & Section \refBad{secMPSintermediateValueTheorem} & 5 &2 \\
&\counterSubTopic && Limits involving $\infty$. & & & & & \\
&& \counterSubSubTopic& Infinite limits, vertical asymptotes. & Chapter 1.6 & Chapter 1.5 & Sections \refBad{secMPSlimitsVerticalAsymptote}, \refBad{secMPShorAndVertAsymptotes} & 6& 3 \\
&& \counterSubSubTopic& Limits at infinity; horizontal asymptotes. & Chapter 1.6 & Chapter 3.4&  Sections \refBad{secMPSlimitsXtoInfty}, \refBad{secMPShorAndVertAsymptotes}& 6& 3 \\\hline 
\counterTopic &&& Inverse functions (Brief Review). & N/A (lectures only)& Chapter 6.1& Section \refBad{secMPSInverseFunctions} & 7& 4 \\\hline
\counterTopic &&& Exponents and Logarithms (Brief review). & N/A (lectures only) & N/A (lectures only)& Section \refBad{secMPSLogarithmsExponentsBasics}& 8& 4 \\\hline\hline
\multicolumn{8}{|c}{Test I}& 5\\\hline\hline
\counterTopic &&& Derivatives. & &&&& \\
&\counterSubTopic && The tangent problems.&  Chapter 2.1& Chapter 1.4& N/A(theory only)&9&5 \\
&\counterSubTopic&& Derivatives and rates of change.& Chapter 2.2& Chapter 2.1& N/A(theory only)& 9& 5 \\
&\counterSubTopic&& The derivative as a function. & Chapter 2.1, 2.2 & Chapter 2.2 & Section \refBad{secMPSderivativesFunGraphsBasics} & 9 & 5\\
&\counterSubTopic&& Basic differentiation formulas, Product rule. & Chapter 2.3, 2.4 & Chapter 2.3 & Section \refBad{secMPSproductQuotientRules} & 10 & 6 \\
&\counterSubTopic&& Derivatives of trigonometric functions. & Chapter 2.3 & Chapter 2.4 & Section \refBad{secMPStrigDerivatives}& 11&6 \\
&\counterSubTopic&& The chain rule. & Chapter 2.5 & Chapter 2.5& Section \refBad{secMPSchainRule} & 12& 7\\
&\counterSubTopic&&\begin{tabular}{l} (Optional) Differentiation f-la proofs \\ from chain and product rules.\end{tabular} & N/A(lectures only)& N/A(lectures only)& N/A(theory only)&  13 & 7  \\
&\counterSubTopic&& Implicit differentiation. & Chapter 2.6 & Chapter 2.6 & Section \refBad{secMPSImplicitDifferentiation}& 14 & 8  \\
&\counterSubTopic&& (Optional) Logarithmic derivatives. & Chapter 2.6 & Chapter 6.3 & Section \refBad{secMPSDerivativeNonConstExponent} & 15 & \skipped \\
&\counterSubTopic&& Related rates.& Chapter 4.2 & Chapter  2.8& Section \refBad{secMPSrelatedRates}& 15 & \skipped \\ \hline
\counterTopic &&& Graphical behavior of functions.& & &&& \\
&\counterSubTopic && The Extreme Value Theorem. & Chapter 3.1 & Chapter 3.1 &N/A(theory only) & 16 & 8\\
&\counterSubTopic && The Mean Value Theorem. & Chapter 3.2& Chapter 3.2& Section \refBad{secMPS-MVT} & 16 & 8 \\
&\counterSubTopic && The closed interval method.& Chapter 4.3& Chapter 3.1& Section \refBad{secMPSclosedInterval} & 17 & 8\\
&\counterSubTopic && The first derivative test. & Chapter 3.3& Chapter 3.3& Section \refBad{secMPSoneVariableMinMax}& 18 & 9 \\
&\counterSubTopic && One variable differentiable function optimization. & Chapter 4.3 & Chapter 3.7& Section \refBad{secMPSoptimization} & 18 & 9 \\
&\counterSubTopic && Concavity.& Chapter 3.4 & Chapter 3.3& Section \refBad{secMPSfunctionGraphSketching}& 18 & 9 \\
&\counterSubTopic && The second derivative test. & Chapter 3.4& Chapter 3.3& Section \refBad{secMPSoneVariableMinMax}& 18 & 9 \\
&\counterSubTopic&& Curve sketching. & Chapter 3.5 & Chapter 3.5& Section \refBad{secMPSfunctionGraphSketching}& 18 & 9 \\\hline
\counterTopic &&& (Optional) Newton's method. & Chapter 4.1 & Chapter 3.8  & N/A (theory only)& 19& \skipped \\\hline\hline
\multicolumn{8}{|c}{Test II} & 10
\\\hline\hline
\counterTopic &&& Linear approximations and differentials.&&&&& \\
&\counterSubTopic && Linear approximations. & Chapter 4.4 & Chapter 2.9 & Section \refBad{secMPSLinearizationAndDifferentials} &19 & 10 \\
&\counterSubTopic && Differentials. & Chapter 4.4 & Chapter 2.9& N/A (theory only)& 19& 10  \\\hline
\counterTopic &&& Integrals.& & & & &\\
&\counterSubTopic && Riemann sums and measuring areas. & Chapter 5.3  & Chapters 4.1, 4.2 & Section \refBad{secMPSRiemannSums}& 21 & 11 \\
&\counterSubTopic && Definite integrals. & && & &\\
&&\counterSubSubTopic& Definite integral definition. & Chapters 5.2, 5.3 & Chapters 4.1, 4.2 & N/A (theory only)& 21 & 11  \\
&&\counterSubSubTopic& Basic properties of integration. & Chapter 5.2 & Chapter 4.2 & N/A (theory only)& 21 & 11\\
&\counterSubTopic && Computing integrals: basics. &&&&&  \\
&&\counterSubSubTopic& Antiderivatives and indefinite integrals & Chapter 5.1 & Chapters 3.9, 4.4 & Sections \refBad{secMPSantiderivatives}, \refBad{secMPSBasicDefiniteIntegrals} & 22 & 11 \\
&&\counterSubSubTopic& The Evaluation Theorem (Fund. Th. Calc., Part 2) & Chapter 5.4 & Chapter 4.3& Section \refBad{secMPSBasicDefiniteIntegrals} & 22& 11\\
&&\counterSubSubTopic & The Fundamental Theorem of Calculus, Part 1. & Chapter 5.4 & Chapter 4.3 & Section \refBad{secMPSFTCpart1} & 23 & 12 \\
&\counterSubTopic && Integration: the substitution rule. &&& Section \refBad{secMPSintegrationSubstitutionRule} & 23 & 12  \\
&&\counterSubSubTopic& The substitution rule in indefinite integrals. & Chapter 6.1& Chapter 4.5 & Section \refBad{secMPSintegrationSubstitutionRuleIndefinite} & 23 &12 \\
&&\counterSubSubTopic& The substitution rule in definite integrals. & Chapter 6.1 & Chapter 4.5 & Section \refBad{secMPSintegrationSubstitutionRuleDefinite} & 23 & 12 \\\hline
\counterTopic&&& First applications of integration.&&&&&\\
&\counterSubSubTopic& & Area between curves. & Chapter 7.1 & Chapter 5.1& Section \refBad{secMPSareaBetweenCurves}& 24& 13 \\
&\counterSubTopic&& Volumes of solids of revolution. &&&Section \refBad{secMPSvolumesSolidsRevolution} && \\
&&\counterSubSubTopic & Volumes of solids of revolution (washer method). & Chapter 7.2& Chapter 5.2 & Section \refBad{secMPSvolumesSolidsRevolutionWashers} & 24 & 13  \\
&&\counterSubSubTopic & Volumes of solids of revolution (cylindrical shell method). & Chapter 7.3& Chapter 5.3 & Section \refBad{secMPSvolumesSolidsRevolutionShells} & 24 & \skipped \\\hline\hline
\multicolumn{8}{|c}{Catch up if behind schedule. Review for the final.} & 14\\\hline
\end{longtable}
\end{landscape}
\end{document}