\documentclass{article}
\input{../homework-problems-UMB}
\renewcommand{\fcProblemRef}{\theenumi.\alph*}

\toggletrue{solutions}
\toggletrue{answers}
%\togglefalse{solutions}

\renewcommand{\course}{Math 130}
\newcommand{\tooeasy}{ \textbf{The problem is too easy to appear on a quiz or test. }}

\begin{comment}
\homeworkStart{on Lecture 1 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/coordinate-systems/homework/plot-the-points-2d-1}
\item \input{../../modules/coordinate-systems/homework/distance-between-points-2d-1}
\input{../../modules/coordinate-systems/homework/distance-between-points-2d-1-solutions}
\item \input{../../modules/coordinate-systems/homework/find-midpoint-2d-1}

\homeworkEnd
\end{comment}

\begin{comment}
\homeworkStart{on Lecture 2 \\ Quiz time will be announced in class.}{}

\item \input{../../modules/coordinate-systems/homework/recognizing-circles-1}
\input{../../modules/coordinate-systems/homework/recognizing-circles-1-solutions}
\item \input{../../modules/coordinate-systems/homework/equation-circle-from-center-and-point-1}

\item \input{../../modules/equation-graph/homework/find-x-y-intercepts-graph-given-circle-1}

\homeworkEnd
\end{comment}
\begin{comment}
\homeworkStart{on Lecture 3 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/lines-2d/homework/line-from-two-points-1}
\item \input{../../modules/lines-2d/homework/find-line-intersection-1}
\input{../../modules/lines-2d/homework/find-line-intersection-1-solutions}

\homeworkEnd
\end{comment}

%\begin{comment}
\homeworkStart{on Lecture 4 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/precalculus/homework/functions-evaluate-simplify}
\item \input{../../modules/precalculus/homework/functions-domains-ranges-1}
\item \input{../../modules/functions-basics/homework/piecewise-linear-formula-from-plot-1}
\item \input{../../modules/functions-basics/homework/even-or-odd-from-formula-1}
\input{../../modules/functions-basics/homework/even-or-odd-from-formula-1-solutions}

%\item \input{../../modules/functions-basics/homework/piecewise-linear-piecewise-circular-formula-from-plot-1}


\homeworkEnd
\end{comment}

\begin{comment}
\homeworkStart{on Lecture 5 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/functions-basics/homework/functions-composing-fractional-linear-1}
\item \input{../../modules/functions-basics/homework/function-composition-domains-and-ranges-1}

\homeworkEnd
\end{comment}
\begin{comment}
\homeworkStart{on Lecture 6 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/function-graph-linear-transformations/homework/use-known-graph-to-sketch-its-linear-transformation-1}
 \input{../../modules/function-graph-linear-transformations/homework/use-known-graph-to-sketch-its-linear-transformation-1-solutions}
\homeworkEnd
\end{comment}

\begin{comment}
\homeworkStart{on Lecture 7 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/quadratic-functions/homework/solve-the-quadratic-equation-1}
\item \input{../../modules/quadratic-functions/homework/plot-the-quadratic-1}
\item \input{../../modules/quadratic-functions/homework/maxima-minima-quadratic-1}
\homeworkEnd
\end{comment}

\begin{comment}
\homeworkStart{on Lecture 8 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/exponential-functions/homework/exponent-simplify}
\input{../../modules/exponential-functions/homework/exponent-simplify-solutions}
\item \input{../../modules/exponential-functions/homework/sketch-exponent-shift-from-known-exponent-1}
\item \input{../../modules/logarithms/homework/e-as-a-limit-and-compound-interest}
\input{../../modules/logarithms/homework/e-as-a-limit-and-compound-interest-solutions}

\homeworkEnd
\end{comment}

\begin{comment}
\homeworkStart{on Lecture 9 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/inverse-functions/homework/inverse-functions3}
\input{../../modules/inverse-functions/homework/inverse-functions3-solutions}
\item \input{../../modules/inverse-functions/homework/inverse-functions4}

\item \textbf{This problem uses material that has not been studied (yet), and will therefore not appear on the quiz.}
\input{../../modules/inverse-functions/homework/inverse-functions2}
\input{../../modules/inverse-functions/homework/inverse-functions2-solutions}

\homeworkEnd
\end{comment}
\begin{comment}
\homeworkStart{on Lecture 10 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/logarithms/homework/logarithms-basic-no-properties-1}
\homeworkEnd
\end{comment}

\begin{comment}
\homeworkStart{on Lecture 11 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/logarithms/homework/logarithms-basic-properties-1}
\input{../../modules/logarithms/homework/logarithms-basic-properties-1-solutions}
\item \input{../../modules/logarithms/homework/logarithms-log-arbitrary-base-from-ln-1}
\input{../../modules/logarithms/homework/logarithms-log-arbitrary-base-from-ln-1-solutions}
\item \input{../../modules/logarithms/homework/logarithms-combine-1}
\input{../../modules/logarithms/homework/logarithms-combine-1-solutions}
\item \input{../../modules/logarithms/homework/logarithms-arctrig-hyperbolic-theoretic-1}

\item \input{../../modules/logarithms/homework/logarithms-equations}
\input{../../modules/logarithms/homework/logarithms-equations-solutions}
\item \input{../../modules/inverse-functions/homework/inverse-functions2}



\homeworkEnd
\end{comment}
\begin{comment}
\homeworkStart{on Lecture 12 \\ Quiz time will be announced in class.}{}

\item \input{../../modules/logarithms/homework/exponential-growth-decay-model-1}
\input{../../modules/logarithms/homework/exponential-growth-decay-model-1-solutions}

\item \input{../../modules/logarithms/homework/logistic-growth-model-1}
\item \input{../../modules/logarithms/homework/logarithmic-model-1}


\homeworkEnd
\end{comment}
\begin{comment}
\homeworkStart{on Lecture 13 \\ Quiz time will be announced in class.}{}

\item \input{../../modules/trigonometry/homework/trigonometry-angle-conversion-degree-to-radian}
\item \input{../../modules/trigonometry/homework/trigonometry-angle-conversion-radian-to-degree}
\item \input{../../modules/trigonometry/homework/arc-length-from-degrees-1}

\homeworkEnd
\end{comment}
\begin{comment}
\homeworkStart{on Lecture 14 \\ Quiz time will be announced in class.}{}

\item \input{../../modules/trigonometry/homework/find-trig-functions-from-right-triangle-1}
\item \input{../../modules/trigonometry/homework/find-exact-trig-function-value-1}
\item \input{../../modules/trigonometry/homework/solve-simplest-trig-equation-1}
\input{../../modules/trigonometry/homework/solve-simplest-trig-equation-1-solutions}
\homeworkEnd
\end{comment}
\begin{comment}
\homeworkStart{on Lecture 15 \\ Quiz time will be announced in class.}{}

\item  \input{../../modules/trigonometry/homework/trig-angle-sum-formulas-to-find-sin-cos-1}
\item  \input{../../modules/trigonometry/homework/angle-sum-formulas-to-compute-cofunction-identities-1}

\item \textbf{ Problems \ref{problemExpresssin^6xviaTrigFunctionsMultipleAngle} and \ref{problemExpresscos^6xviaTrigFunctionsMultipleAngle} are considered challenge problems and will not be tested/quizzed upon.}
\input{../../modules/trigonometry/homework/power-reducing-1}

\item \input{../../modules/trigonometry/homework/sum-to-product-formulas-to-solve-trig-equations-1}

\homeworkEnd
\end{comment}

\begin{comment}
\homeworkStart{on Lecture 16 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/trigonometry/homework/trigonometry-identities}
\homeworkEnd
\end{comment}
%\begin{comment}
\homeworkStart{on Lecture 17 \\ Quiz time will be announced in class.}{}
\item \input{../../modules/inverse-trig/homework/inverse-trig-evaluated-on-trig-2}
\input{../../modules/inverse-trig/homework/inverse-trig-evaluated-on-trig-2-solutions}
\item \input{../../modules/inverse-trig/homework/trig-evaluated-on-inverse-trig-2}
\input{../../modules/inverse-trig/homework/trig-evaluated-on-inverse-trig-2-solutions}
\item \input{../../modules/inverse-trig/homework/trig-evaluated-on-inverse-trig-1}
\input{../../modules/inverse-trig/homework/trig-evaluated-on-inverse-trig-1-solutions}

\homeworkEnd
\end{comment}


