% begin module exponential-def-various-approaches
\begin{frame}
\frametitle{Exponents overview}
\begin{itemize}
\item<1-> For integer $x$, we know how to compute $a^{x}$ as a function of $a$. 
\item<2-> However, how do we compute $f(x)=a^x$ when $x$ is not an integer?
\item<3-> We need to go back to the definition of $a^x$ (when $x$ is not an integer).
\item<4-> In what follows we give/recall the Calc I definition of exponent.
\item<5-> Then we give an alternative (but of course equivalent) second definition of exponent. The second definition is studied in detail by the end of Calculus II. 
\item<6-> We discuss pros and cons of the two definitions. In a nutshell: the Calc I definition is easier to motivate, the Calc II definition is easier to compute with.
\end{itemize}
\end{frame}
\begin{frame}
\frametitle{Exponent definition using limits (approach I)}
\begin{itemize}
\item<1-> For integer $p$ we know to compute $a^p$.
\item<2-> Therefore for integer $q$ we know to compute $a^{\frac{1}{q}}= \sqrt[q]{a}=\max\{x|\text{~for~which~} x^q\leq a\}$.
\item<3-> Therefore we know to compute $a^{\frac{p}{q}}$ for all rational $\frac{p}{q}$.
\item<4-> We can then define
\[
a^x = \lim\limits_{\substack{y \to x \\ y\text{-rational}}} a^y 
\]
\item<5-> Not computationally effective. It is not how computers compute.
\item<6-> However is the easiest approach. $a^{x+y}=a^xa^y$ is easiest to prove (follows directly from the $\varepsilon, \delta$-definition of $\lim$).
\item<7->\alertNoH{7-}{This is the definition assumed in Calculus I.}
\end{itemize}
\end{frame}
\begin{frame}
\frametitle{Exponent definition using series (approach II)}
\begin{itemize}
\item<1-> The following formula (studied at end of Calc II) can be used as alternative definition.
\[
e^{x}=\alertNoH{4}{\sum_{n=0}^{\infty}} \frac{x^n}{\alertNoH{2}{n!}}= 1+ x+\frac{x^2}{2!}+\frac{x^3}{3!}+\dots + \frac{x^{n}}{\alertNoH{2}{n!}}+\dots
\]
\uncover<2->{\alertNoH{2}{Here $n!=1\cdot 2\cdot 3\cdot\dots \cdot(n-1)\cdot n$ and is read ``$n$ factorial''. }}
\item<3-> For $|x|<1$ define 
\[
\ln (1+x)=\alertNoH{4}{\sum_{n=1}^{\infty}} (-1)^{n+1}\frac{x^n}{n}=  x-\frac{x^2}{2}+\frac{x^3}{3}-\dots + \frac{(-1)^{n+1}x^{n}}{n}+\dots
\]
\uncover<4->{\alertNoH{4}{Infinite sum studied in Calc II.}} 
\item<5-> For arbitrary $a>0$ define $a^x$ as $a^x=e^{x\ln a}$. 
\item<6-> Disadvantage: more difficult to prove $e^{x+y}= e^{x} e^y$ and $e^{\ln(1+x)}=1+x$, proof done in Calculus II.
\item<7-> This is how computers compute $e^x$ and $a^x$.
\end{itemize}
\end{frame}

% end module exponential-def-various-approaches
